---
title: "Study 2 — Wave4+ Clean Pipeline (Attention + Hardcoded Longstring)"
output:
  html_document:
    df_print: paged
    code_folding: hide
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

# 0) Config

## This sets up options for knitting the document into a PDF or HTML.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = TRUE, warning = TRUE)
options(cli.num_colors = 1, crayon.enabled = FALSE)
```

## File path (localized into the git branch)

- **Edit the `path`** below to your CSV export.
- Column names are **hard-coded** to your Wave4+ survey. If your export uses different labels, adjust in the **Column Map** section.
- This file creates:  
  - `FirstItem` … `SeventhItem` (1–4: position of selected choice)  
  - `longstring_run` via `{careless}` across those seven items  
  - `exclude_reason` with `"attention_fail"` and `"longstring_outlier"`  
  - `dat_final` after all exclusions

```{r config}
# ---- PATH ----
path <- r"(Dissertation+-+Study+2+-+Wave4+_October+6,+2025_15.23.csv)"
```

# 1) Packages

```{r packages}
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
library(pacman)
pacman::p_load(readr, dplyr, stringr, ggplot2, broom, effectsize, careless)
```

# 2) Load (Extract from CSV)

```{r load}
raw0 <- readr::read_csv(
  path,
  col_types = cols(.default = col_character()),
  show_col_types = FALSE
)

# Drop the first two header-ish rows (labels + ImportId JSON)
stopifnot(nrow(raw0) >= 3)
dat <- raw0[-c(1,2), , drop = FALSE]

# Keep finished only (exact header "Finished")
dat <- dat %>% filter(Finished %in% c("True","true","1"))

# Coerce duration and Score
dat[["Duration (in seconds)"]] <- suppressWarnings(as.numeric(dat[["Duration (in seconds)"]]))
dat[["SC0"]]                   <- suppressWarnings(as.numeric(dat[["SC0"]]))
dat$Score <- dat[["SC0"]]

stopifnot(nrow(dat) > 0)
stopifnot(!all(is.na(dat$Score)))
```

# 3) Derive factors (Version → Hierarchy & chartjunk)

```{r factors}
stopifnot("Version" %in% names(dat))
dat$Version <- stringr::str_squish(dat$Version)

chart <- stringr::str_match(dat$Version, "(High|Medium|Low)Chartjunk$")[,2]
hier  <- stringr::str_remove(dat$Version, "(High|Medium|Low)Chartjunk$")

dat$chartjunk <- factor(chart, levels = c("Low","Medium","High"), ordered = TRUE)
dat$Hierarchy <- factor(
  dplyr::recode(hier,
    "NoHierarchy"         = "NoHierarchy",
    "FloorplanHierarchy"  = "FloorplanHierarchy",
    "RoomHierarchy"       = "RoomHierarchy",
    "ImportanceHierarchy" = "ImportanceHierarchy",
    .default = NA_character_
  ),
  levels = c("NoHierarchy","FloorplanHierarchy","RoomHierarchy","ImportanceHierarchy")
)

## Recoding dat$Hierarchy
dat$Hierarchy <- as.character(dat$Hierarchy)
dat$Hierarchy[dat$Hierarchy == "NoHierarchy"] <- "None"
dat$Hierarchy[dat$Hierarchy == "FloorplanHierarchy"] <- "Floorplan"
dat$Hierarchy[dat$Hierarchy == "RoomHierarchy"] <- "Room"
dat$Hierarchy[dat$Hierarchy == "ImportanceHierarchy"] <- "Importance"

## Reordering dat$Hierarchy
dat$Hierarchy <- factor(dat$Hierarchy,
  levels = c("Importance", "None", "Room", "Floorplan")
)

## Reordering dat$Hierarchy
dat$Hierarchy <- factor(dat$Hierarchy,
  levels = c("Importance", "None", "Room", "Floorplan")
)
```

# 4) Column Map (hard-coded)

These are the exact headers expected in the Wave4+ export. 
For copying users, adjust here if your CSV labels differ.

```{r colmap}
# === Column Map (Wave4+ CSV using QIDs) ===
# AN display order (tokens like Q1|Q9|...|MetaInfo)
COL_AN      <- "Gistblock_DO"

# Seven items (Q1 & Q18 have per-row randomized choice order)
COL_Q1      <- "Q1"
COL_Q1_DO   <- "Q1_DO"

COL_Q9      <- "Q9"
COL_Q10     <- "Q10"
COL_Q14     <- "Q14"
COL_Q17     <- "Q17"

COL_Q18     <- "Q18"
COL_Q18_DO  <- "Q18_DO"

# Attention item (also one of the seven)
COL_Q20     <- "Q20"

# Fixed (non-DO) choice orders for 4-option questions
# (coerce source columns to character so match() works even if CSV parsed numerics)
dat[[COL_Q10]] <- as.character(dat[[COL_Q10]])
dat[[COL_Q14]] <- as.character(dat[[COL_Q14]])
dat[[COL_Q17]] <- as.character(dat[[COL_Q17]])
dat[[COL_Q9 ]] <- trimws(as.character(dat[[COL_Q9 ]]))
dat[[COL_Q20]] <- trimws(as.character(dat[[COL_Q20]]))

CHO_Q9  <- c("North","South","East","West")
CHO_Q10 <- c("0","1","2","3")
CHO_Q14 <- c("1","2","3","4")
CHO_Q17 <- c("0","1","2","3")
CHO_Q20 <- c("Smart home","IT infrastructure","Hospital bed display","Stock ticker")

req_cols <- c(COL_AN,COL_Q1,COL_Q1_DO,COL_Q9,COL_Q10,COL_Q14,COL_Q17,COL_Q18,COL_Q18_DO,COL_Q20)
if (!all(req_cols %in% names(dat))) {
  stop("Missing columns: ", paste(setdiff(req_cols, names(dat)), collapse = ", "))
}

```

# 5) Attention filter (Q20 must contain "smart home")

## Step one of careless responding removal -- a standard manipulation check.

```{r attention}
keep_idx <- grepl("smart[[:space:]]*home", dat[[COL_Q20]], ignore.case = TRUE)
keep_idx[is.na(keep_idx)] <- FALSE

dat$exclude_reason <- ifelse(keep_idx, NA_character_, "attention_fail")
dat_attn <- dat[keep_idx, , drop = FALSE]

message(sprintf("Attention filter kept %d / %d rows.", sum(keep_idx, na.rm = TRUE), nrow(dat)))
```

# 6) Build FirstItem…SeventhItem (1–4 positions)

Rules:
- Parse `COL_AN` (pipe-separated tokens) → drop `MetaInfo` → take the first seven tokens as the **seen order**.
- Map each token to one of `{Q1, Q9, Q10, Q14, Q17, Q18, Q20}`.
- For Q1 & Q18, compute the selected **position** from the per-row “- Display Order” columns.
- For fixed-order items, compute the position via `match(selected, CHO_*)`.

```{r first_to_seventh}
# 6a) Parse AN into seven tokens per row
ORD_LIST <- strsplit(dat_attn[[COL_AN]], "\\|", perl = TRUE)
ORD_LIST <- lapply(ORD_LIST, function(v) {
  v <- v[!grepl("^\\s*MetaInfo\\s*$", v, ignore.case = TRUE)]
  if (length(v) >= 7) v[1:7] else c(v, rep(NA_character_, 7 - length(v)))
})

# 6b) Selected positions per item (1..4)

# Q1 and Q18: position derived from their row-specific Display Order
POS_Q1 <- as.integer(mapply(function(ans, do) {
  if (is.na(ans) || is.na(do)) return(NA_integer_)
  ch <- strsplit(do, "\\|", perl = TRUE)[[1]]
  match(ans, ch)
}, dat_attn[[COL_Q1]], dat_attn[[COL_Q1_DO]]))

POS_Q18 <- as.integer(mapply(function(ans, do) {
  if (is.na(ans) || is.na(do)) return(NA_integer_)
  ch <- strsplit(do, "\\|", perl = TRUE)[[1]]
  match(ans, ch)
}, dat_attn[[COL_Q18]], dat_attn[[COL_Q18_DO]]))

# Fixed-order items
POS_Q9  <- as.integer(match(dat_attn[[COL_Q9]],  CHO_Q9))
POS_Q10 <- as.integer(match(dat_attn[[COL_Q10]], CHO_Q10))
POS_Q14 <- as.integer(match(dat_attn[[COL_Q14]], CHO_Q14))
POS_Q17 <- as.integer(match(dat_attn[[COL_Q17]], CHO_Q17))
POS_Q20 <- as.integer(match(dat_attn[[COL_Q20]], CHO_Q20))

# Mapping token → vector of positions in fixed order c(Q1,Q9,Q10,Q14,Q17,Q18,Q20)
QKEYS <- c("Q1","Q9","Q10","Q14","Q17","Q18","Q20")

# Helper inline (no separate functions): index into the packed vector by token
pick_by_token <- function(tok, p1,p9,p10,p14,p17,p18,p20) {
  idx <- match(tok, QKEYS)
  c(p1,p9,p10,p14,p17,p18,p20)[idx]
}

FirstItem   <- as.integer(mapply(function(ord, p1,p9,p10,p14,p17,p18,p20) pick_by_token(ord[1], p1,p9,p10,p14,p17,p18,p20),
                                ORD_LIST, POS_Q1, POS_Q9, POS_Q10, POS_Q14, POS_Q17, POS_Q18, POS_Q20))
SecondItem  <- as.integer(mapply(function(ord, p1,p9,p10,p14,p17,p18,p20) pick_by_token(ord[2], p1,p9,p10,p14,p17,p18,p20),
                                ORD_LIST, POS_Q1, POS_Q9, POS_Q10, POS_Q14, POS_Q17, POS_Q18, POS_Q20))
ThirdItem   <- as.integer(mapply(function(ord, p1,p9,p10,p14,p17,p18,p20) pick_by_token(ord[3], p1,p9,p10,p14,p17,p18,p20),
                                ORD_LIST, POS_Q1, POS_Q9, POS_Q10, POS_Q14, POS_Q17, POS_Q18, POS_Q20))
FourthItem  <- as.integer(mapply(function(ord, p1,p9,p10,p14,p17,p18,p20) pick_by_token(ord[4], p1,p9,p10,p14,p17,p18,p20),
                                ORD_LIST, POS_Q1, POS_Q9, POS_Q10, POS_Q14, POS_Q17, POS_Q18, POS_Q20))
FifthItem   <- as.integer(mapply(function(ord, p1,p9,p10,p14,p17,p18,p20) pick_by_token(ord[5], p1,p9,p10,p14,p17,p18,p20),
                                ORD_LIST, POS_Q1, POS_Q9, POS_Q10, POS_Q14, POS_Q17, POS_Q18, POS_Q20))
SixthItem   <- as.integer(mapply(function(ord, p1,p9,p10,p14,p17,p18,p20) pick_by_token(ord[6], p1,p9,p10,p14,p17,p18,p20),
                                ORD_LIST, POS_Q1, POS_Q9, POS_Q10, POS_Q14, POS_Q17, POS_Q18, POS_Q20))
SeventhItem <- as.integer(mapply(function(ord, p1,p9,p10,p14,p17,p18,p20) pick_by_token(ord[7], p1,p9,p10,p14,p17,p18,p20),
                                ORD_LIST, POS_Q1, POS_Q9, POS_Q10, POS_Q14, POS_Q17, POS_Q18, POS_Q20))

dat_attn$FirstItem   <- FirstItem
dat_attn$SecondItem  <- SecondItem
dat_attn$ThirdItem   <- ThirdItem
dat_attn$FourthItem  <- FourthItem
dat_attn$FifthItem   <- FifthItem
dat_attn$SixthItem   <- SixthItem
dat_attn$SeventhItem <- SeventhItem
```

# 7) Longstring via `{careless}`

We compute longest run across the seven 1–4 codes. To avoid NA runs, we replace missing entries with **unique per-column sentinels** so consecutive NAs never match.

## Step two of careless responding
  1) Determine the longstring value
  2) Compute Tukey's Fence
  3) Remove rows where longstring > fence

```{r longstring}
ls_mat <- dat_attn[, c("FirstItem","SecondItem","ThirdItem","FourthItem","FifthItem","SixthItem","SeventhItem")]

# Clean blanks → NA_integer_
for (j in seq_along(ls_mat)) {
  x <- ls_mat[[j]]
  x[x %in% c("", "NA")] <- NA
  ls_mat[[j]] <- suppressWarnings(as.integer(x))
}

# Replace NA with unique per-column codes (breaks NA runs)
for (j in seq_along(ls_mat)) {
  idx <- is.na(ls_mat[[j]])
  if (any(idx)) ls_mat[[j]][idx] <- -j
}

dat_attn$longstring_run <- as.integer(careless::longstring(ls_mat))

# Tukey fence to flag outliers; set fixed cutoff if preferred
q3  <- stats::quantile(dat_attn$longstring_run, 0.75, na.rm = TRUE)
iqr <- stats::IQR(dat_attn$longstring_run, na.rm = TRUE)
thr <- q3 + 1.5 * iqr
flag <- !is.na(dat_attn$longstring_run) & dat_attn$longstring_run > thr

dat_attn$exclude_reason <- ifelse(
  is.na(dat_attn$exclude_reason) & flag, "longstring_outlier",
  ifelse(!is.na(dat_attn$exclude_reason) & flag,
         paste(dat_attn$exclude_reason, "+ longstring_outlier"),
         dat_attn$exclude_reason)
)

dat_final <- dat_attn[!flag, , drop = FALSE]

cat("Longstring on First..Seventh: threshold >", round(thr, 2),
    "| removed", sum(flag, na.rm = TRUE), "rows.\n")
```

# 8) Quick EDA (optional)

```{r eda, fig.width=7, fig.height=4}
# Score ~ Hierarchy
ggplot(dat_final, aes(Hierarchy, Score)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = .12, outlier.shape = NA) +
  geom_jitter(width = .12, height = 0, alpha = .25, size = 1) +
  labs(y = "Score", x = "Hierarchy") +
  theme_minimal(base_size = 12)

# Score ~ chartjunk
ggplot(dat_final, aes(chartjunk, Score)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = .12, outlier.shape = NA) +
  geom_jitter(width = .12, height = 0, alpha = .25, size = 1) +
  labs(x = "Chartjunk", y = "Score") +
  ylim (c(1,7))+
  theme_minimal(base_size = 12)
```

# 9) Sanity tables

```{r sanity}
table(dat_attn$exclude_reason, useNA = "ifany")
summary(dat_attn$longstring_run)

# Peek at the First..Seventh encodings
head(dat_attn[, c("FirstItem","SecondItem","ThirdItem","FourthItem","FifthItem","SixthItem","SeventhItem")], 10)
```

# 10) Session info

```{r session}
sessionInfo()
```


```{r}
# ---- ANCOVA: adjust for dashboard use (Never vs Otherwise) ----
suppressPackageStartupMessages({
  p_load(car)
  p_load(emmeans)
  p_load(effectsize)
  p_load(ggplot2)
})

# 1) Recode usage to binary factor (ref = Otherwise)
dat_final$use_never <- ifelse(tolower(trimws(dat_final$DataDashboardUsage)) == "never",
                              "Never", "Otherwise")
dat_final$use_never <- factor(dat_final$use_never, levels = c("Otherwise","Never"))

# (Optional) sanity check
table(dat_final$use_never, useNA = "ifany")

# 2) Fit ANCOVA (binary covariate)
m_anc <- lm(Score ~ Hierarchy * chartjunk + use_never, data = dat_final)

# 3) Type-III tests with sum contrasts
old_contr <- options(contrasts = c("contr.sum","contr.poly"))
anc_tests <- car::Anova(m_anc, type = 3)
options(old_contr)

anc_tests

# 4) Effect sizes (partial eta^2)
eta_squared(anc_tests, partial = TRUE)

# 5) Adjusted means (EMMs) for Hierarchy x chartjunk
#    weights = "proportional" averages over use_never by its sample proportions.
#    If you prefer equal weighting across use_never levels, set weights = "equal".
emm_anc <- emmeans(m_anc, ~ Hierarchy * chartjunk, weights = "proportional")
confint(emm_anc)

# 6) Plot adjusted means with 95% CIs
emm_df <- as.data.frame(confint(emm_anc))
ggplot(emm_df, aes(Hierarchy, emmean, group = chartjunk, color = chartjunk)) +
  geom_line(position = position_dodge(width = .2)) +
  geom_point(position = position_dodge(width = .2), size = 2.5) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = .12, position = position_dodge(width = .2)) +
  labs(x = "Hierarchy", y = "Adjusted mean Score (ANCOVA)",
       color = "Chartjunk") +
  theme_minimal(base_size = 12)

# 7) (Optional) follow-ups
# Pairwise simple effects of chartjunk within each Hierarchy (Tukey-adjusted)
pairs(emmeans(m_anc, ~ chartjunk | Hierarchy, weights = "proportional"),
      adjust = "tukey")

# Pairwise simple effects of Hierarchy within each chartjunk
pairs(emmeans(m_anc, ~ Hierarchy | chartjunk, weights = "proportional"),
      adjust = "tukey")

# 8) (Optional) model diagnostics
par(mfrow = c(1,2))
plot(m_anc, which = 1)  # Residuals vs Fitted
plot(m_anc, which = 2)  # Normal Q-Q
par(mfrow = c(1,1))

```

```{r}
library(pacman)
p_load(ggthemes)
# ---- sjPlot visuals (ANCOVA) ----
pacman::p_load(sjPlot, sjlabelled)

# Optional: consistent look + quiet color codes in console
theme_set(sjPlot::theme_sjplot())
options(cli.num_colors = 1, crayon.enabled = FALSE)

# 1) Interaction: Hierarchy × chartjunk (averages over use_never by sample mix)
p_int <- sjPlot::plot_model(
  m_anc,
  type  = "int",
  terms = c("Hierarchy", "chartjunk"),   # x, group
  ci.lvl = 0.95
) + labs(x = "Hierarchy", y = "Adjusted mean Score") +
    guides(linetype = "none", shape = "none")
p_int

# 2) Predicted means, explicitly showing both use_never levels
# terms = c(x, group, facet)
p_pred <- sjPlot::plot_model(
  m_anc,
  type  = "pred",
  terms = c("Hierarchy", "chartjunk", "use_never")
) + labs(x = "Hierarchy", y = "Adjusted mean Score", color = "Chartjunk") +
    guides(linetype = "none", shape = "none")
p_pred

# 3) Coefficient forest (fixed effects with 95% CI)
p_coef <- sjPlot::plot_model(
  m_anc,
  type = "est",          # estimated coefficients
  show.values = TRUE,
  value.offset = .2
) + labs(x = NULL, y = "Coefficient (±95% CI)")
p_coef

# 4) Model table (nice HTML output in the notebook)
sjPlot::tab_model(
  m_anc,
  dv.labels = "Score (ANCOVA)",
  show.ci = TRUE,
  show.std = TRUE,
  p.style = "stars",
  string.pred = "Predictor",
  string.est  = "b"
) + theme_tufte()

```

```{r}
# 1) Interaction: Hierarchy × chartjunk (averages over use_never by sample mix)
p_Hierarchy <- sjPlot::plot_model(
  m_anc,
  type  = "eff",
  terms = c("Hierarchy"),   # x, group
  ci.lvl = 0.95
) + labs(x = "Hierarchy", y = "Adjusted mean Score") +
    guides(linetype = "none", shape = "none") # + theme_tufte()
p_Hierarchy
```


```{r}
m_mod <- lm(Score ~ Hierarchy * chartjunk * use_never, data = dat_final)
car::Anova(m_mod, type = 3)

```
```{r}
m_aov <- lm(Score ~ Hierarchy * chartjunk, data = dat_final)
m_anc <- lm(Score ~ Hierarchy * chartjunk + use_never, data = dat_final)

anova(m_aov, m_anc)   # nested F-test
AIC(m_aov, m_anc); BIC(m_aov, m_anc)

```
```{r}
# --- Final 2×3 ANOVA (no covariate) ---
p_load(car); p_load(emmeans); p_load(effectsize); p_load(sjPlot)

m_aov <- lm(Score ~ Hierarchy * chartjunk, data = dat_final)

# Type III with sum contrasts (reporting-friendly)
oldc <- options(contrasts = c("contr.sum","contr.poly"))
aov_tests <- car::Anova(m_aov, type = 3)
options(oldc)

aov_tests
effectsize::eta_squared(aov_tests, partial = TRUE)

# EMMs + pairwise (Tukey)
emm <- emmeans(m_aov, ~ Hierarchy * chartjunk)
pairs(emmeans(m_aov, ~ chartjunk | Hierarchy), adjust = "tukey")
pairs(emmeans(m_aov, ~ Hierarchy | chartjunk), adjust = "tukey")

# Plot (report-friendly)
theme_set(sjPlot::theme_sjplot())
sjPlot::plot_model(m_aov, type = "int",
                   terms = c("Hierarchy","chartjunk"), ci.lvl = 0.95) +
  labs(x = "Hierarchy", y = "Estimated mean Score") +
  # Use colorblind-friendly palette
  scale_color_brewer(palette = "RdYlBu") 

# (Optional) basic assumptions
car::leveneTest(Score ~ interaction(Hierarchy, chartjunk), data = dat_final)
par(mfrow = c(1,2)); plot(m_aov, which = 1); plot(m_aov, which = 2); par(mfrow = c(1,1))

```

```{r}
library(sjPlot)
library(ggplot2)

theme_set(theme_sjplot())

# Okabe–Ito contrast-safe diverging palette
okabe_ito_div <- c(
  "Low" = "#E69F00",   # orange
  "Medium" = "#999999",# gray
  "High" = "#0072B2"   # deep blue
)

p <- plot_model(m_aov, type = "int",
                terms = c("Hierarchy", "chartjunk"),
                ci.lvl = 0.95) +
  labs(x = "Hierarchy", y = "Estimated mean Score") +
  scale_color_manual(values = okabe_ito_div, aesthetics = c("colour", "fill")) +
  theme(
    text = element_text(size = 12),
    legend.position = "top",
    legend.title = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line = element_line(color = "grey30"),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA)
  ) +
  # Enhance line/point visibility
  guides(color = guide_legend(override.aes = list(size = 3))) # +
  # geom_line(size = 1.1, alpha = 0.9) +
  # geom_point(size = 3, stroke = 0.8)

p

```


```{r}
# 1) Omnibus 2×3 ANOVA + effect sizes
options(contrasts = c("contr.sum","contr.poly"))
m_aov <- lm(Score ~ Hierarchy * chartjunk, data = dat_final)
aov_tab <- car::Anova(m_aov, type = 3); aov_tab
effectsize::eta_squared(aov_tab, partial = TRUE)

# 2) Planned linear trend for chartjunk (it’s ordered)
# overall (balanced over Hierarchy) and within each Hierarchy
p_load(emmeans)
emm_cj   <- emmeans(m_aov, ~ chartjunk, weights = "equal")
contrast(emm_cj, "poly")              # look at the Linear row

emm_cj_h <- emmeans(m_aov, ~ chartjunk | Hierarchy, weights = "equal")
contrast(emm_cj_h, "poly")            # Linear per-Hierarchy

# 3) Simple effects of chartjunk within each Hierarchy (Tukey)
pairs(emm_cj_h, adjust = "tukey")

# (Optional) visualize the monotone look you’re seeing
emm_df <- as.data.frame(confint(emm_cj_h))
ggplot(emm_df, aes(chartjunk, emmean, group = Hierarchy, color = Hierarchy)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = .12, position = position_dodge(.2)) +
  labs(x = "Chartjunk", y = "Estimated mean Score (balanced)") +
  theme_minimal(base_size = 12)

```
```{r}
# ensure ordered levels
dat_final$chartjunk <- factor(dat_final$chartjunk,
                               levels = c("Low","Medium","High"),
                               ordered = TRUE)

# contrasts: sum for Hierarchy (Type III friendly), polynomial for chartjunk
contrasts(dat_final$Hierarchy) <- contr.sum(4)
contrasts(dat_final$chartjunk)  <- contr.poly(3)

# refit the 2×3 model
m_aov <- lm(Score ~ Hierarchy * chartjunk, data = dat_final)

# overall Type III tests
car::Anova(m_aov, type = 3)

```

```{r}
p_load(emmeans)

# overall linear trend for chartjunk (balanced over Hierarchy)
emm_cj <- emmeans(m_aov, ~ chartjunk, weights = "equal")
contrast(emm_cj, "poly")   # check the "Linear" row

# linear trend within each Hierarchy
emm_cj_h <- emmeans(m_aov, ~ chartjunk | Hierarchy, weights = "equal")
contrast(emm_cj_h, "poly") # Linear per Hierarchy level

```

```{r}
# ---- linear contrast at a fixed Hierarchy level -----------------------------
# deps
p_load(emmeans)
p_load(ggplot2)
p_load(sjPlot)  # for theme_sjplot()

plot_linear_contrast_at <- function(
  m,
  focal,                     # character: the factor you want the linear contrast on (e.g., "cond")
  moderator = "Hierarchy",   # character: the factor to hold constant
  at_level = NULL,           # which level of moderator to hold fixed (e.g., "Manager")
  contrast = c("poly","custom"),  # "poly" for polynomial linear trend, or "custom"
  custom_weights = NULL,     # numeric vector if contrast = "custom" (length = #levels of focal)
  type = "response",         # EMM scale: "response" often nicer for GLMs
  emm_weights = "equal"      # how to average over other factors if present; "equal" is common
) {
  contrast <- match.arg(contrast)

  # discover default at_level if not provided
  mf <- tryCatch(model.frame(m), error = function(e) NULL)
  if (is.null(at_level)) {
    if (!is.null(mf) && moderator %in% names(mf) && is.factor(mf[[moderator]])) {
      at_level <- levels(mf[[moderator]])[1]
    } else {
      stop("Please provide 'at_level' (couldn't infer a factor level for ", moderator, ").")
    }
  }

  # build a reference grid with Hierarchy held constant
  rg <- ref_grid(m, at = setNames(list(at_level), moderator), type = type)

  # EMMs for the focal factor at that Hierarchy level
  emm <- emmeans(rg, as.formula(paste0("~ ", focal)), weights = emm_weights)

  # Linear contrast
  lin <- if (contrast == "poly") {
    # "poly" gives linear, quadratic, ...; take the linear component only
    contrast(emm, "poly")["linear"]
  } else {
    # custom weights you define in the order of levels(emm)
    if (is.null(custom_weights))
      stop("Provide 'custom_weights' when contrast = 'custom'.")
    if (length(custom_weights) != nrow(as.data.frame(emm)))
      stop("Length of 'custom_weights' must equal the number of levels in '", focal, "'.")
    contrast(emm, list(linear = custom_weights))
  }

  lin_sum <- summary(lin, infer = TRUE)
  tval <- lin_sum$t.ratio
  pval <- lin_sum$p.value

  # data frame for plotting
  emm_df <- as.data.frame(emm)
  emm_df$.xnum <- as.numeric(emm_df[[focal]])

  p <- ggplot(emm_df,
              aes(x = .xnum,
                  y = emmean,
                  ymin = lower.CL, ymax = upper.CL)) +
    geom_point(size = 3) +
    geom_line(group = 1) +
    geom_errorbar(width = 0.1) +
    geom_smooth(method = "lm", se = FALSE) +
    scale_x_continuous(breaks = emm_df$.xnum,
                       labels = emm_df[[focal]]) +
    labs(
      x = focal,
      y = paste("Estimated marginal mean (", type, " scale)", sep = ""),
      title = paste0("Linear contrast across ", focal,
                     " (", moderator, " = ", at_level, ")"),
      subtitle = paste0("t = ", round(tval, 2),
                        ", p = ", format.pval(pval, digits = 3, eps = 1e-4))
    ) +
    theme_sjplot()

  list(plot = p, emm = emm, linear_contrast = lin)
}
# -----------------------------------------------------------------------------

# EXAMPLE USAGE:
# pacman::p_load(lme4)
# m <- lmer(y ~ cond * Hierarchy + (1|id), data = d)  # or lm()/glm()
# out <- plot_linear_contrast_at(m, focal = "cond", moderator = "Hierarchy",
#                                at_level = "Manager", contrast = "poly",
#                                type = "response", emm_weights = "equal")
# out$plot
# # summary(out$linear_contrast)  # full stats table

```

```{r}
# --- Linear contrasts for chartjunk while holding Hierarchy constant ----

# pick which fitted model you want to use
# (no covariate 2×3)     or     (ANCOVA with use_never)
mod <- m_aov              # or: mod <- m_anc

# how emmeans averages over other factors not shown in the formula
# "equal" = simple average of cells; "proportional" = weighted by sample sizes
wts <- "equal"            # or: wts <- "proportional"

# Uses your existing m_aov (or m_anc), dat_final, and plot_linear_contrast_at()
emm_cj_h <- emmeans(mod, ~ chartjunk | Hierarchy, weights = wts)
emm_df   <- as.data.frame(confint(emm_cj_h))
emm_df$xnum <- as.numeric(emm_df$chartjunk)

ggplot(emm_df, aes(x = xnum, y = emmean)) +
  facet_wrap(~ Hierarchy) +
  geom_point(size = 2.5) +
  geom_line(group = 1) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = .12) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_continuous(breaks = seq_along(levels(dat_final$chartjunk)),
                     labels = levels(dat_final$chartjunk)) +
  labs(x = "Chartjunk", y = "Estimated marginal mean") +
  theme_sjplot()


```
```{r}
# --- 3-point line charts for chartjunk: raw vs. adjusted ---------------------
pacman::p_load(dplyr, ggplot2, emmeans, tibble, patchwork, sjPlot)

# Ensure ordered factor
dat_final$chartjunk <- factor(dat_final$chartjunk,
                              levels = c("Low","Medium","High"),
                              ordered = TRUE)

# If needed, fit (you already have m_aov from above)
if (!exists("m_aov")) {
  m_aov <- lm(Score ~ Hierarchy * chartjunk, data = dat_final)
}

# 1) RAW means (ignore Hierarchy entirely)
raw_df <- dat_final %>%
  group_by(chartjunk) %>%
  summarize(
    n = sum(!is.na(Score)),
    mean = mean(Score, na.rm = TRUE),
    sd = sd(Score, na.rm = TRUE),
    se = sd / sqrt(n),
    tcrit = qt(.975, df = pmax(n - 1, 1)),
    lower = mean - tcrit * se,
    upper = mean + tcrit * se,
    .groups = "drop"
  )

p_raw <- ggplot(raw_df, aes(x = chartjunk, y = mean, group = 1)) +
  geom_line() +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = .12) +
  labs(x = "Chartjunk", y = "Score", title = "Raw means (ignoring Hierarchy)") +
  theme_sjplot()

# 2) ADJUSTED means (EMMs) for chartjunk, balanced over Hierarchy
emm_cj_eq <- emmeans(m_aov, ~ chartjunk, weights = "equal")
emm_df <- as.data.frame(confint(emm_cj_eq))

p_adj <- ggplot(emm_df, aes(x = chartjunk, y = emmean, group = 1)) +
  geom_line() +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = .12) +
  labs(x = "Chartjunk", y = "Estimated marginal mean",
       title = "EMMs balanced over Hierarchy") +
  theme_sjplot()

# Show side by side
p_raw | p_adj
# ------------------------------------------------------------------------------

# If you want the adjusted version to use sample-mix instead of equal weighting:
# emm_cj_prop <- emmeans(m_aov, ~ chartjunk, weights = "proportional")
# as.data.frame(confint(emm_cj_prop))

```
```{r}
## Recoding dat$Hierarchy
dat$Hierarchy <- as.character(dat$Hierarchy)
dat$Hierarchy[dat$Hierarchy == "NoHierarchy"] <- "None"
dat$Hierarchy[dat$Hierarchy == "FloorplanHierarchy"] <- "Floorplan"
dat$Hierarchy[dat$Hierarchy == "RoomHierarchy"] <- "Room"
dat$Hierarchy[dat$Hierarchy == "ImportanceHierarchy"] <- "Importance"
```

# 11) sjPlot of ANOVA

```{r}
library(sjPlot)
library(ggplot2)

p <- sjPlot::plot_model(
  m_aov,
  type  = "eff",
  terms = c("chartjunk", "Hierarchy"),
  ci.lvl = 0.95
) +
  labs(x = "Chartjunk", y = "Estimated mean Score") + ylim(c(2.4,4.0))

# Add lines between levels for each Hierarchy group
p + geom_line() -> q

# Save as 1080p PNG with white background
ggsave(
  filename = "chartjunk_Hierarchy_1080p.png",
  plot     = q,
  width    = 1920,
  height   = 1080,
  units    = "px",
  bg       = "white"
)

```

```{r}
library(sjPlot)
library(ggplot2)

p <- sjPlot::plot_model(
  m_aov,
  type  = "eff",
  terms = c("chartjunk", "Hierarchy"),
  ci.lvl = 0.95
) +
  labs(x = "Chartjunk", y = "Estimated mean Score") +
  ylim(c(2.4, 4.0))

q <- p +
  geom_line() +
  facet_wrap(~group, nrow = 1) +     # one facet per Hierarchy level, single row
  theme_bw(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "none",         # legend is redundant once faceted
    panel.spacing.x  = unit(0.8, "lines")  # ← increase gap between panels
    )

q


# Save as 1080p PNG with white background
ggsave(
  filename = "chartjunk_Hierarchy_faceted_1080p.png",
  plot     = q,
  width    = 1920,
  height   = 1080,
  units    = "px",
  bg       = "white"
)
```


```{r}
p <- sjPlot::plot_model(
  m_aov,
  type  = "eff",
  terms = c("chartjunk"),
  ci.lvl = 0.95
) +
  labs(x = "Hierarchy", y = "Estimated mean Score") + ylim(c(2.4,4))

# Add lines between levels for each Hierarchy group
p + geom_line()
```

```{r}



p <- sjPlot::plot_model(
  m_aov,
  type  = "eff",
  terms = c("Hierarchy"),
  ci.lvl = 0.95
) +
  labs(x = "Hierarchy", y = "Estimated mean Score") +
  ylim(c(2.4, 4)) 

p
```

```{r}
if (!requireNamespace("pacman", quietly = TRUE)) {
  install.packages("pacman")
}

pacman::p_load(
  sjPlot,
  ggplot2,
  ggh4x,
  scales
)

Hierarchy_cols <- c(
  Importance = "#E41A1C",
  None       = "#377EB8",
  Room       = "#4DAF4A",
  Floorplan  = "#984EA3"
)

p <- sjPlot::plot_model(
  m_aov,
  type  = "eff",
  terms = c("chartjunk", "Hierarchy"),
  ci.lvl = 0.95
) +
  labs(x = "Chartjunk", y = "Estimated mean Score") +
  ylim(c(2.4, 4.0)) +
  scale_colour_manual(values = Hierarchy_cols) +
  theme_bw(base_size = 12)

q <- p +
  geom_line() +
  ggh4x::facet_wrap2(
    ~group,
    nrow = 1,
    strip = ggh4x::strip_themed(
      background_x = ggh4x::elem_list_rect(
        fill   = alpha(Hierarchy_cols, 0.18),
        colour = "grey40"
      ),
      text_x = ggh4x::elem_list_text(colour = "black")
    )
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position  = "none",
    strip.background = element_blank(),
        panel.spacing.x  = unit(0.8, "lines")  # ← increase gap between panels
  )

q

```

```{r}
ggsave(
  filename = "chartjunk_Hierarchy_facets_1080p.png",
  plot     = q,
  width    = 1920,
  height   = 1080,
  units    = "px",
  bg       = "white"
)

```

# 12) Effect sizes from ANOVA

```{r}
# assuming anova_model is your 2x2 ANOVA (aov/lm or afex object)
p_load(effectsize)

round_df <- function(x, k = 3) {
  x <- as.data.frame(x)
  num <- vapply(x, is.numeric, logical(1))
  x[num] <- lapply(x[num], round, k)
  x
}

es_omega  <- omega_squared(m_aov)
es_eta_p  <- eta_squared(m_aov, partial = TRUE)
es_eta_g  <- eta_squared(m_aov, generalized = TRUE)

round_df(es_omega)
round_df(es_eta_p)
round_df(es_eta_g)
```

```{r}
summary(m_aov)
```

