---
title: "Study 2 — Wave4+ Clean Pipeline (Attention + Hardcoded Longstring)"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = TRUE, warning = TRUE)
options(cli.num_colors = 1, crayon.enabled = FALSE)
```

# 0) Config

- **Edit the `path`** below to your CSV export.
- Column names are **hard-coded** to your Wave4+ survey. If your export uses different labels, adjust in the **Column Map** section.
- This file creates:  
  - `FirstItem` … `SeventhItem` (1–4: position of selected choice)  
  - `longstring_run` via `{careless}` across those seven items  
  - `exclude_reason` with `"attention_fail"` and `"longstring_outlier"`  
  - `dat_final` after all exclusions

```{r config}
# ---- PATH ----
path <- r"(D:\Dropbox\USD\Dissertation\Collab folder\Data\Dissertation+-+Study+2+-+Wave4+_October+6,+2025_15.23.csv)"
```

# 1) Packages

```{r packages}
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
pacman::p_load(readr, dplyr, stringr, ggplot2, broom, effectsize, careless)
```

# 2) Load

```{r load}
raw0 <- readr::read_csv(
  path,
  col_types = cols(.default = col_character()),
  show_col_types = FALSE
)

# Drop the first two header-ish rows (labels + ImportId JSON)
stopifnot(nrow(raw0) >= 3)
dat <- raw0[-c(1,2), , drop = FALSE]

# Keep finished only (exact header "Finished")
dat <- dat %>% filter(Finished %in% c("True","true","1"))

# Coerce duration and score
dat[["Duration (in seconds)"]] <- suppressWarnings(as.numeric(dat[["Duration (in seconds)"]]))
dat[["SC0"]]                   <- suppressWarnings(as.numeric(dat[["SC0"]]))
dat$score <- dat[["SC0"]]

stopifnot(nrow(dat) > 0)
stopifnot(!all(is.na(dat$score)))
```

# 3) Derive factors (Version → hierarchy & chartjunk)

```{r factors}
stopifnot("Version" %in% names(dat))
dat$Version <- stringr::str_squish(dat$Version)

chart <- stringr::str_match(dat$Version, "(High|Medium|Low)Chartjunk$")[,2]
hier  <- stringr::str_remove(dat$Version, "(High|Medium|Low)Chartjunk$")

dat$chartjunk <- factor(chart, levels = c("Low","Medium","High"), ordered = TRUE)
dat$hierarchy <- factor(
  dplyr::recode(hier,
    "NoHierarchy"         = "NoHierarchy",
    "FloorplanHierarchy"  = "FloorplanHierarchy",
    "RoomHierarchy"       = "RoomHierarchy",
    "ImportanceHierarchy" = "ImportanceHierarchy",
    .default = NA_character_
  ),
  levels = c("NoHierarchy","FloorplanHierarchy","RoomHierarchy","ImportanceHierarchy")
)
```

# 4) Column Map (hard-coded)

These are the exact headers expected in your Wave4+ export. Adjust here if your CSV labels differ.

```{r colmap}
# === Column Map (Wave4+ CSV using QIDs) ===
# AN display order (tokens like Q1|Q9|...|MetaInfo)
COL_AN      <- "Gistblock_DO"

# Seven items (Q1 & Q18 have per-row randomized choice order)
COL_Q1      <- "Q1"
COL_Q1_DO   <- "Q1_DO"

COL_Q9      <- "Q9"
COL_Q10     <- "Q10"
COL_Q14     <- "Q14"
COL_Q17     <- "Q17"

COL_Q18     <- "Q18"
COL_Q18_DO  <- "Q18_DO"

# Attention item (also one of the seven)
COL_Q20     <- "Q20"

# Fixed (non-DO) choice orders for 4-option questions
# (coerce source columns to character so match() works even if CSV parsed numerics)
dat[[COL_Q10]] <- as.character(dat[[COL_Q10]])
dat[[COL_Q14]] <- as.character(dat[[COL_Q14]])
dat[[COL_Q17]] <- as.character(dat[[COL_Q17]])
dat[[COL_Q9 ]] <- trimws(as.character(dat[[COL_Q9 ]]))
dat[[COL_Q20]] <- trimws(as.character(dat[[COL_Q20]]))

CHO_Q9  <- c("North","South","East","West")
CHO_Q10 <- c("0","1","2","3")
CHO_Q14 <- c("1","2","3","4")
CHO_Q17 <- c("0","1","2","3")
CHO_Q20 <- c("Smart home","IT infrastructure","Hospital bed display","Stock ticker")

req_cols <- c(COL_AN,COL_Q1,COL_Q1_DO,COL_Q9,COL_Q10,COL_Q14,COL_Q17,COL_Q18,COL_Q18_DO,COL_Q20)
if (!all(req_cols %in% names(dat))) {
  stop("Missing columns: ", paste(setdiff(req_cols, names(dat)), collapse = ", "))
}

```

# 5) Attention filter (Q20 must contain "smart home")

```{r attention}
keep_idx <- grepl("smart[[:space:]]*home", dat[[COL_Q20]], ignore.case = TRUE)
keep_idx[is.na(keep_idx)] <- FALSE

dat$exclude_reason <- ifelse(keep_idx, NA_character_, "attention_fail")
dat_attn <- dat[keep_idx, , drop = FALSE]

message(sprintf("Attention filter kept %d / %d rows.", sum(keep_idx, na.rm = TRUE), nrow(dat)))
```

# 6) Build FirstItem…SeventhItem (1–4 positions)

Rules:
- Parse `COL_AN` (pipe-separated tokens) → drop `MetaInfo` → take the first seven tokens as the **seen order**.
- Map each token to one of `{Q1, Q9, Q10, Q14, Q17, Q18, Q20}`.
- For Q1 & Q18, compute the selected **position** from the per-row “- Display Order” columns.
- For fixed-order items, compute the position via `match(selected, CHO_*)`.

```{r first_to_seventh}
# 6a) Parse AN into seven tokens per row
ORD_LIST <- strsplit(dat_attn[[COL_AN]], "\\|", perl = TRUE)
ORD_LIST <- lapply(ORD_LIST, function(v) {
  v <- v[!grepl("^\\s*MetaInfo\\s*$", v, ignore.case = TRUE)]
  if (length(v) >= 7) v[1:7] else c(v, rep(NA_character_, 7 - length(v)))
})

# 6b) Selected positions per item (1..4)

# Q1 and Q18: position derived from their row-specific Display Order
POS_Q1 <- as.integer(mapply(function(ans, do) {
  if (is.na(ans) || is.na(do)) return(NA_integer_)
  ch <- strsplit(do, "\\|", perl = TRUE)[[1]]
  match(ans, ch)
}, dat_attn[[COL_Q1]], dat_attn[[COL_Q1_DO]]))

POS_Q18 <- as.integer(mapply(function(ans, do) {
  if (is.na(ans) || is.na(do)) return(NA_integer_)
  ch <- strsplit(do, "\\|", perl = TRUE)[[1]]
  match(ans, ch)
}, dat_attn[[COL_Q18]], dat_attn[[COL_Q18_DO]]))

# Fixed-order items
POS_Q9  <- as.integer(match(dat_attn[[COL_Q9]],  CHO_Q9))
POS_Q10 <- as.integer(match(dat_attn[[COL_Q10]], CHO_Q10))
POS_Q14 <- as.integer(match(dat_attn[[COL_Q14]], CHO_Q14))
POS_Q17 <- as.integer(match(dat_attn[[COL_Q17]], CHO_Q17))
POS_Q20 <- as.integer(match(dat_attn[[COL_Q20]], CHO_Q20))

# Mapping token → vector of positions in fixed order c(Q1,Q9,Q10,Q14,Q17,Q18,Q20)
QKEYS <- c("Q1","Q9","Q10","Q14","Q17","Q18","Q20")

# Helper inline (no separate functions): index into the packed vector by token
pick_by_token <- function(tok, p1,p9,p10,p14,p17,p18,p20) {
  idx <- match(tok, QKEYS)
  c(p1,p9,p10,p14,p17,p18,p20)[idx]
}

FirstItem   <- as.integer(mapply(function(ord, p1,p9,p10,p14,p17,p18,p20) pick_by_token(ord[1], p1,p9,p10,p14,p17,p18,p20),
                                ORD_LIST, POS_Q1, POS_Q9, POS_Q10, POS_Q14, POS_Q17, POS_Q18, POS_Q20))
SecondItem  <- as.integer(mapply(function(ord, p1,p9,p10,p14,p17,p18,p20) pick_by_token(ord[2], p1,p9,p10,p14,p17,p18,p20),
                                ORD_LIST, POS_Q1, POS_Q9, POS_Q10, POS_Q14, POS_Q17, POS_Q18, POS_Q20))
ThirdItem   <- as.integer(mapply(function(ord, p1,p9,p10,p14,p17,p18,p20) pick_by_token(ord[3], p1,p9,p10,p14,p17,p18,p20),
                                ORD_LIST, POS_Q1, POS_Q9, POS_Q10, POS_Q14, POS_Q17, POS_Q18, POS_Q20))
FourthItem  <- as.integer(mapply(function(ord, p1,p9,p10,p14,p17,p18,p20) pick_by_token(ord[4], p1,p9,p10,p14,p17,p18,p20),
                                ORD_LIST, POS_Q1, POS_Q9, POS_Q10, POS_Q14, POS_Q17, POS_Q18, POS_Q20))
FifthItem   <- as.integer(mapply(function(ord, p1,p9,p10,p14,p17,p18,p20) pick_by_token(ord[5], p1,p9,p10,p14,p17,p18,p20),
                                ORD_LIST, POS_Q1, POS_Q9, POS_Q10, POS_Q14, POS_Q17, POS_Q18, POS_Q20))
SixthItem   <- as.integer(mapply(function(ord, p1,p9,p10,p14,p17,p18,p20) pick_by_token(ord[6], p1,p9,p10,p14,p17,p18,p20),
                                ORD_LIST, POS_Q1, POS_Q9, POS_Q10, POS_Q14, POS_Q17, POS_Q18, POS_Q20))
SeventhItem <- as.integer(mapply(function(ord, p1,p9,p10,p14,p17,p18,p20) pick_by_token(ord[7], p1,p9,p10,p14,p17,p18,p20),
                                ORD_LIST, POS_Q1, POS_Q9, POS_Q10, POS_Q14, POS_Q17, POS_Q18, POS_Q20))

dat_attn$FirstItem   <- FirstItem
dat_attn$SecondItem  <- SecondItem
dat_attn$ThirdItem   <- ThirdItem
dat_attn$FourthItem  <- FourthItem
dat_attn$FifthItem   <- FifthItem
dat_attn$SixthItem   <- SixthItem
dat_attn$SeventhItem <- SeventhItem
```

# 7) Longstring via `{careless}`

We compute longest run across the seven 1–4 codes. To avoid NA runs, we replace missing entries with **unique per-column sentinels** so consecutive NAs never match.

```{r longstring}
ls_mat <- dat_attn[, c("FirstItem","SecondItem","ThirdItem","FourthItem","FifthItem","SixthItem","SeventhItem")]

# Clean blanks → NA_integer_
for (j in seq_along(ls_mat)) {
  x <- ls_mat[[j]]
  x[x %in% c("", "NA")] <- NA
  ls_mat[[j]] <- suppressWarnings(as.integer(x))
}

# Replace NA with unique per-column codes (breaks NA runs)
for (j in seq_along(ls_mat)) {
  idx <- is.na(ls_mat[[j]])
  if (any(idx)) ls_mat[[j]][idx] <- -j
}

dat_attn$longstring_run <- as.integer(careless::longstring(ls_mat))

# Tukey fence to flag outliers; set fixed cutoff if preferred
q3  <- stats::quantile(dat_attn$longstring_run, 0.75, na.rm = TRUE)
iqr <- stats::IQR(dat_attn$longstring_run, na.rm = TRUE)
thr <- q3 + 1.5 * iqr
flag <- !is.na(dat_attn$longstring_run) & dat_attn$longstring_run > thr

dat_attn$exclude_reason <- ifelse(
  is.na(dat_attn$exclude_reason) & flag, "longstring_outlier",
  ifelse(!is.na(dat_attn$exclude_reason) & flag,
         paste(dat_attn$exclude_reason, "+ longstring_outlier"),
         dat_attn$exclude_reason)
)

dat_final <- dat_attn[!flag, , drop = FALSE]

cat("Longstring on First..Seventh: threshold >", round(thr, 2),
    "| removed", sum(flag, na.rm = TRUE), "rows.\n")
```

# 8) Quick EDA (optional)

```{r eda, fig.width=7, fig.height=4}
# Score ~ hierarchy
ggplot(dat_final, aes(hierarchy, score)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = .12, outlier.shape = NA) +
  geom_jitter(width = .12, height = 0, alpha = .25, size = 1) +
  labs(y = "Score", x = "Hierarchy") +
  theme_minimal(base_size = 12)

# Score ~ chartjunk
ggplot(dat_final, aes(chartjunk, score)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = .12, outlier.shape = NA) +
  geom_jitter(width = .12, height = 0, alpha = .25, size = 1) +
  labs(x = "Chartjunk", y = "Score") +
  theme_minimal(base_size = 12)
```

# 9) Sanity tables

```{r sanity}
table(dat_attn$exclude_reason, useNA = "ifany")
summary(dat_attn$longstring_run)

# Peek at the First..Seventh encodings
head(dat_attn[, c("FirstItem","SecondItem","ThirdItem","FourthItem","FifthItem","SixthItem","SeventhItem")], 10)
```

# 10) Session info

```{r session}
sessionInfo()
```


```{r}
# ---- ANCOVA: adjust for dashboard use (Never vs Otherwise) ----
suppressPackageStartupMessages({
  library(car)
  library(emmeans)
  library(effectsize)
  library(ggplot2)
})

# 1) Recode usage to binary factor (ref = Otherwise)
dat_final$use_never <- ifelse(tolower(trimws(dat_final$DataDashboardUsage)) == "never",
                              "Never", "Otherwise")
dat_final$use_never <- factor(dat_final$use_never, levels = c("Otherwise","Never"))

# (Optional) sanity check
table(dat_final$use_never, useNA = "ifany")

# 2) Fit ANCOVA (binary covariate)
m_anc <- lm(score ~ hierarchy * chartjunk + use_never, data = dat_final)

# 3) Type-III tests with sum contrasts
old_contr <- options(contrasts = c("contr.sum","contr.poly"))
anc_tests <- car::Anova(m_anc, type = 3)
options(old_contr)

anc_tests

# 4) Effect sizes (partial eta^2)
eta_squared(anc_tests, partial = TRUE)

# 5) Adjusted means (EMMs) for hierarchy x chartjunk
#    weights = "proportional" averages over use_never by its sample proportions.
#    If you prefer equal weighting across use_never levels, set weights = "equal".
emm_anc <- emmeans(m_anc, ~ hierarchy * chartjunk, weights = "proportional")
confint(emm_anc)

# 6) Plot adjusted means with 95% CIs
emm_df <- as.data.frame(confint(emm_anc))
ggplot(emm_df, aes(hierarchy, emmean, group = chartjunk, color = chartjunk)) +
  geom_line(position = position_dodge(width = .2)) +
  geom_point(position = position_dodge(width = .2), size = 2.5) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = .12, position = position_dodge(width = .2)) +
  labs(x = "Hierarchy", y = "Adjusted mean score (ANCOVA)",
       color = "Chartjunk") +
  theme_minimal(base_size = 12)

# 7) (Optional) follow-ups
# Pairwise simple effects of chartjunk within each hierarchy (Tukey-adjusted)
pairs(emmeans(m_anc, ~ chartjunk | hierarchy, weights = "proportional"),
      adjust = "tukey")

# Pairwise simple effects of hierarchy within each chartjunk
pairs(emmeans(m_anc, ~ hierarchy | chartjunk, weights = "proportional"),
      adjust = "tukey")

# 8) (Optional) model diagnostics
par(mfrow = c(1,2))
plot(m_anc, which = 1)  # Residuals vs Fitted
plot(m_anc, which = 2)  # Normal Q-Q
par(mfrow = c(1,1))

```

```{r}
# ---- sjPlot visuals (ANCOVA) ----
pacman::p_load(sjPlot, sjlabelled)

# Optional: consistent look + quiet color codes in console
theme_set(sjPlot::theme_sjplot())
options(cli.num_colors = 1, crayon.enabled = FALSE)

# 1) Interaction: hierarchy × chartjunk (averages over use_never by sample mix)
p_int <- sjPlot::plot_model(
  m_anc,
  type  = "int",
  terms = c("hierarchy", "chartjunk"),   # x, group
  ci.lvl = 0.95
) + labs(x = "Hierarchy", y = "Adjusted mean score") +
    guides(linetype = "none", shape = "none")
p_int

# 2) Predicted means, explicitly showing both use_never levels
# terms = c(x, group, facet)
p_pred <- sjPlot::plot_model(
  m_anc,
  type  = "pred",
  terms = c("hierarchy", "chartjunk", "use_never")
) + labs(x = "Hierarchy", y = "Adjusted mean score", color = "Chartjunk") +
    guides(linetype = "none", shape = "none")
p_pred

# 3) Coefficient forest (fixed effects with 95% CI)
p_coef <- sjPlot::plot_model(
  m_anc,
  type = "est",          # estimated coefficients
  show.values = TRUE,
  value.offset = .2
) + labs(x = NULL, y = "Coefficient (±95% CI)")
p_coef

# 4) Model table (nice HTML output in the notebook)
sjPlot::tab_model(
  m_anc,
  dv.labels = "Score (ANCOVA)",
  show.ci = TRUE,
  show.std = TRUE,
  p.style = "stars",
  string.pred = "Predictor",
  string.est  = "b"
)

```

```{r}
m_mod <- lm(score ~ hierarchy * chartjunk * use_never, data = dat_final)
car::Anova(m_mod, type = 3)

```
```{r}
m_aov <- lm(score ~ hierarchy * chartjunk, data = dat_final)
m_anc <- lm(score ~ hierarchy * chartjunk + use_never, data = dat_final)

anova(m_aov, m_anc)   # nested F-test
AIC(m_aov, m_anc); BIC(m_aov, m_anc)

```
```{r}
# --- Final 2×3 ANOVA (no covariate) ---
library(car); library(emmeans); library(effectsize); library(sjPlot)

m_aov <- lm(score ~ hierarchy * chartjunk, data = dat_final)

# Type III with sum contrasts (reporting-friendly)
oldc <- options(contrasts = c("contr.sum","contr.poly"))
aov_tests <- car::Anova(m_aov, type = 3)
options(oldc)

aov_tests
effectsize::eta_squared(aov_tests, partial = TRUE)

# EMMs + pairwise (Tukey)
emm <- emmeans(m_aov, ~ hierarchy * chartjunk)
pairs(emmeans(m_aov, ~ chartjunk | hierarchy), adjust = "tukey")
pairs(emmeans(m_aov, ~ hierarchy | chartjunk), adjust = "tukey")

# Plot (report-friendly)
theme_set(sjPlot::theme_sjplot())
sjPlot::plot_model(m_aov, type = "int",
                   terms = c("hierarchy","chartjunk"), ci.lvl = 0.95) +
  labs(x = "Hierarchy", y = "Estimated mean score") +
  guides(linetype = "none", shape = "none")

# (Optional) basic assumptions
car::leveneTest(score ~ interaction(hierarchy, chartjunk), data = dat_final)
par(mfrow = c(1,2)); plot(m_aov, which = 1); plot(m_aov, which = 2); par(mfrow = c(1,1))

```

```{r}
# 1) Omnibus 2×3 ANOVA + effect sizes
options(contrasts = c("contr.sum","contr.poly"))
m_aov <- lm(score ~ hierarchy * chartjunk, data = dat_final)
aov_tab <- car::Anova(m_aov, type = 3); aov_tab
effectsize::eta_squared(aov_tab, partial = TRUE)

# 2) Planned linear trend for chartjunk (it’s ordered)
# overall (balanced over hierarchy) and within each hierarchy
library(emmeans)
emm_cj   <- emmeans(m_aov, ~ chartjunk, weights = "equal")
contrast(emm_cj, "poly")              # look at the Linear row

emm_cj_h <- emmeans(m_aov, ~ chartjunk | hierarchy, weights = "equal")
contrast(emm_cj_h, "poly")            # Linear per-hierarchy

# 3) Simple effects of chartjunk within each hierarchy (Tukey)
pairs(emm_cj_h, adjust = "tukey")

# (Optional) visualize the monotone look you’re seeing
emm_df <- as.data.frame(confint(emm_cj_h))
ggplot(emm_df, aes(chartjunk, emmean, group = hierarchy, color = hierarchy)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = .12, position = position_dodge(.2)) +
  labs(x = "Chartjunk", y = "Estimated mean score (balanced)") +
  theme_minimal(base_size = 12)

```
```{r}
# ensure ordered levels
dat_final$chartjunk <- factor(dat_final$chartjunk,
                               levels = c("Low","Medium","High"),
                               ordered = TRUE)

# contrasts: sum for hierarchy (Type III friendly), polynomial for chartjunk
contrasts(dat_final$hierarchy) <- contr.sum(4)
contrasts(dat_final$chartjunk)  <- contr.poly(3)

# refit the 2×3 model
m_aov <- lm(score ~ hierarchy * chartjunk, data = dat_final)

# overall Type III tests
car::Anova(m_aov, type = 3)

```

```{r}
library(emmeans)

# overall linear trend for chartjunk (balanced over hierarchy)
emm_cj <- emmeans(m_aov, ~ chartjunk, weights = "equal")
contrast(emm_cj, "poly")   # check the "Linear" row

# linear trend within each hierarchy
emm_cj_h <- emmeans(m_aov, ~ chartjunk | hierarchy, weights = "equal")
contrast(emm_cj_h, "poly") # Linear per hierarchy level

```

