---
title: "Study 2 — Positions (QSF keys), Attention Filter, SurveyWave-Aware Longstring, and Resolution Covariates"
output: html_document

---

```{r setup, message=FALSE, warning=FALSE}
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
pacman::p_load(readr, dplyr, stringr, janitor, ggplot2, jsonlite, car, effectsize, sjPlot, vcd, rstatix, tidyr, purrr, tibble)

# a handy %||% for defaults
`%||%` <- function(a,b) if (is.null(a) || length(a)==0) b else a

# ---- User paths (edit if needed) ----
# Windows raw-string literal style is supported in R >= 4.0 as r"(...)"
csv_path <- r"(D:/Dropbox/USD/Dissertation/Collab folder/Data/Dissertation+-+Study+2_September+30,+2025_10.01.csv)"
qsf_path <- r"(D:/Dropbox/USD/Dissertation/Collab folder/Data/Dissertation_-_Study_2.qsf)"

stopifnot(file.exists(csv_path), file.exists(qsf_path))

# ---- Load CSV (skip first meta row, drop the Qualtrics "ImportId" row if present) ----
raw <- readr::read_csv(csv_path, skip = 1, col_types = readr::cols(.default = readr::col_character()), show_col_types = FALSE)

dat <- raw %>%
  dplyr::filter(!dplyr::if_any(dplyr::everything(), ~ grepl("ImportId", .x, fixed = TRUE)))

dat_orig <- dat  # keep a copy
```

# Map QSF → CSV, compute per-question positions
We derive answer **keys** from the QSF and compute respondent-specific **positions** based on each row's _Display Order_ (or the default order if the DO column is absent).

```{r qsf-positions, message=FALSE, warning=FALSE}
# ---- Read QSF & extract Question metadata for Q1..Q20 ----
qsf <- jsonlite::fromJSON(qsf_path, simplifyVector = FALSE)
sq  <- Filter(function(el) identical(el$Element, "SQ"), qsf$SurveyElements)

get_tag   <- function(p) p$DataExportTag %||% NA_character_
get_qid   <- function(p) p$QuestionID    %||% NA_character_
get_text  <- function(p) p$QuestionText  %||% NA_character_
get_rand  <- function(p) { ra <- p$Randomization %||% list(); if (isTRUE(ra$Advanced) || isTRUE(ra$RandomizeAnswers) || !is.null(ra$Type)) TRUE else FALSE }
get_choice_order <- function(p) { ord <- p$ChoiceOrder %||% names(p$Choices); unname(as.character(ord)) }
get_choices_tbl <- function(p) {
  ch <- p$Choices %||% list()
  keys <- names(ch) %||% character(0)
  if (!length(keys)) return(tibble::tibble(key=character(0), text=character(0), recode=character(0)))
  rv <- p$RecodeValues %||% list()
  recode_vec <- vapply(keys, function(k) as.character(rv[[k]] %||% NA_character_), character(1))
  tibble::tibble(
    key    = as.character(keys),
    text   = vapply(ch, function(x) x$Display %||% "", character(1)),
    recode = recode_vec
  )
}

q_meta <- lapply(sq, function(el) {
  p <- el$Payload
  tibble::tibble(tag=get_tag(p), qid=get_qid(p), text=get_text(p), randomize=get_rand(p),
                 choice_order_default=list(get_choice_order(p)),
                 choices=list(get_choices_tbl(p)))
}) %>%
  dplyr::bind_rows() %>%
  dplyr::filter(!is.na(tag), grepl("^Q([1-9]|1[0-9]|20)$", tag)) %>%
  dplyr::arrange(factor(tag, levels = paste0("Q", 1:20)))

# ---- Match QSF question text to CSV columns ----
find_csv_col <- function(q_text, cols){
  # 1) exact match
  m <- which(cols == q_text)[1]; if (!is.na(m)) return(cols[m])
  # 2) starts with question text
  m <- which(startsWith(cols, q_text))[1]; if (!is.na(m)) return(cols[m])
  # 3) fallback NA
  NA_character_
}
cols <- names(dat)

q_meta <- q_meta %>% dplyr::mutate(
  csv_col       = vapply(text, find_csv_col, character(1), cols = cols),
  csv_col_order = ifelse(is.na(csv_col), NA_character_, paste0(csv_col, " - Display Order"))
)

# ---- Safe regex helpers (POSIX classes; no need for perl=TRUE) ----
answer_to_key <- function(answer_val, choices){
  if (is.na(answer_val) || !length(choices$key)) return(NA_character_)
  ans <- as.character(answer_val)
  # 1) exact text match
  idx <- match(ans, choices$text);    if (!is.na(idx)) return(choices$key[idx])
  # 2) recode match (when export stores numeric recode)
  idx <- match(ans, choices$recode);  if (!is.na(idx)) return(choices$key[idx])
  # 3) forms like "1 - Choice text": extract leading digits
  lead <- regmatches(ans, regexpr("^[[:space:]]*[0-9]+", ans))
  if (length(lead) && nchar(lead)) {
    lead <- sub("^[[:space:]]*([0-9]+).*", "\\1", lead)
    idx <- match(lead, choices$key)
    if (!is.na(idx)) return(choices$key[idx])
  }
  # 4) if the stored value is already a key
  if (ans %in% choices$key) return(ans)
  NA_character_
}

order_to_keys <- function(order_str, choices, default_keys){
  if (is.na(order_str) || !nzchar(order_str)) return(default_keys)
  # detect delimiter
  delim <- if (grepl("\\|", order_str)) "\\|" else if (grepl(";", order_str)) ";" else ","
  toks <- trimws(unlist(strsplit(order_str, delim)))
  if (!length(toks)) return(default_keys)
  # A) tokens are already keys
  if (all(toks %in% choices$key)) return(toks)
  # B) tokens are full texts
  if (all(toks %in% choices$text)) return(choices$key[match(toks, choices$text)])
  # C) tokens like "1 - Choice text": extract leading digits
  lead <- sub("^[[:space:]]*([0-9]+).*", "\\1", toks)
  if (all(lead %in% choices$key)) return(lead)
  # fallback
  default_keys
}

# ---- Build per-QID position columns, then coalesce to tag-level Q*_position ----
pos_map <- list()

for (i in seq_len(nrow(q_meta))) {
  meta <- q_meta[i, ]
  qcol <- meta$csv_col
  if (is.na(qcol)) next
  odcol    <- meta$csv_col_order
  choices  <- meta$choices[[1]]
  def_keys <- meta$choice_order_default[[1]]
  qid      <- meta$qid
  tag      <- meta$tag
  pos_name <- paste0(tag, "_", qid, "_position")

  # map the stored answer (text/recode/key) to the QSF choice key
  ans_keys <- vapply(dat[[qcol]], answer_to_key, character(1), choices = choices)

  # parse each row's Display Order into a vector of keys (list-column)
  ord_keys <- if (!is.na(odcol) && odcol %in% names(dat)) {
    lapply(dat[[odcol]], order_to_keys, choices = choices, default_keys = def_keys)
  } else {
    rep(list(def_keys), nrow(dat))
  }

  # position = index of chosen key in that row's order vector
  pos <- mapply(function(k, ord) {
    if (is.na(k) || !length(ord)) return(NA_integer_)
    p <- match(k, ord)
    if (is.na(p)) NA_integer_ else as.integer(p)
  }, ans_keys, ord_keys, SIMPLIFY = TRUE, USE.NAMES = FALSE)

  dat[[pos_name]] <- as.integer(pos)
  pos_map[[pos_name]] <- pos_name
}

# ---- Coalesce per-tag across QIDs -> single Q*_position ----
tags <- unique(q_meta$tag)
for (tg in tags) {
  cols_tg <- grep(paste0("^", tg, "_.*_position$"), names(dat), value = TRUE)
  if (length(cols_tg)) {
    dat[[paste0(tg, "_position")]] <- purrr::reduce(dat[cols_tg], dplyr::coalesce)
    # (optional) make sure it's integer
    dat[[paste0(tg, "_position")]] <- as.integer(dat[[paste0(tg, "_position")]])
  }
}


# Keep vector of the coalesced Q*_position columns that exist
pos_cols_Q1_Q20 <- paste0("Q", 1:20, "_position")
pos_cols_Q1_Q20 <- pos_cols_Q1_Q20[pos_cols_Q1_Q20 %in% names(dat)]
length(pos_cols_Q1_Q20)
```

# Attention check — keep only “Smart Home”
We drop respondents who don’t report **Smart Home** (case-insensitive, allows optional whitespace).

```{r attention, message=FALSE, warning=FALSE}
# Find the attention question column
attn_col <- grep("^What type of dashboard did you see\\??$", names(dat),
                 ignore.case = TRUE, value = TRUE)
if (length(attn_col) > 1) attn_col <- attn_col[1]

if (length(attn_col) == 0) {
  message("Attention question not found—keeping all rows.")
  dat_attn <- dat
  dat_attn$exclude_reason <- NA_character_
} else {
  keep_idx <- grepl("smart[[:space:]]*home", dat[[attn_col]], ignore.case = TRUE)
  keep_idx[is.na(keep_idx)] <- FALSE  # treat NAs as failures
  dat$exclude_reason <- ifelse(keep_idx, NA_character_, "attention_fail")
  dat_attn <- dat[keep_idx, , drop = FALSE]
  message(sprintf("Attention filter kept %d / %d rows.", sum(keep_idx, na.rm = TRUE), nrow(dat)))
}
```

# Longstring (position-based), **SurveyWave-aware** exclusion of Q3 for Waves 1–2
If `SurveyWave` is 1 or 2, we **exclude Q3** from the longstring per respondent; if 3, we include it. Threshold is the upper IQR fence (*Q3 + 1.5 × IQR*).

```{r longstring, message=FALSE, warning=FALSE}
# find SurveyWave column
wave_col <- grep("^\\s*survey\\s*wave\\s*$", names(dat_attn), ignore.case = TRUE, value = TRUE)
wave_num <- if (length(wave_col)) suppressWarnings(readr::parse_number(dat_attn[[wave_col[1]]])) else rep(NA_real_, nrow(dat_attn))

# which position columns exist (coalesced per question)
present_pos <- pos_cols_Q1_Q20[pos_cols_Q1_Q20 %in% names(dat_attn)]

# identify Q3 position column(s)
q3_pos <- if ("Q3_position" %in% names(dat_attn)) "Q3_position" else {
  grep("^Q3(_.*)?_position$", names(dat_attn), value = TRUE)
}
q3_pos <- intersect(q3_pos, present_pos)

# helper: longest run length (ignores NA)
longest_run <- function(x) {
  x <- x[!is.na(x)]
  if (!length(x)) return(0L)
  r <- rle(x)
  max(r$lengths)
}

# build per-row set of columns to use based on SurveyWave
use_cols_by_row <- lapply(seq_len(nrow(dat_attn)), function(i) {
  cols <- present_pos
  if (length(q3_pos) && !is.na(wave_num[i]) && wave_num[i] %in% c(1, 2)) {
    cols <- setdiff(cols, q3_pos)  # exclude Q3 for Waves 1–2
  }
  cols
})

# compute row-wise longest run
ls_run <- vapply(seq_along(use_cols_by_row), function(i) {
  cols <- use_cols_by_row[[i]]
  if (!length(cols)) return(0L)
  v <- as.integer(dat_attn[i, cols, drop = FALSE])
  longest_run(v)
}, integer(1))

# IQR fence threshold and flag
q3p  <- stats::quantile(ls_run, 0.75, na.rm = TRUE)
iqr  <- stats::IQR(ls_run, na.rm = TRUE)
thr  <- q3p + 1.5 * iqr
flag <- ls_run > thr

dat_attn$longstring_run <- ls_run
dat_attn$exclude_reason <- ifelse(is.na(dat_attn$exclude_reason) & flag, "longstring_outlier",
                           ifelse(!is.na(dat_attn$exclude_reason) & flag,
                                  paste(dat_attn$exclude_reason, "+ longstring_outlier"),
                                  dat_attn$exclude_reason))
dat_final <- dat_attn[!flag, , drop = FALSE]

cat("SurveyWave-aware longstring: threshold >", round(thr, 2),
    "| removed", sum(flag, na.rm = TRUE), "rows.\n")
```

# Covariates — screen **resolution** (height & area)
We parse a `resolution`-like column to derive `res_height` and `res_area`.

```{r covariates, message=FALSE, warning=FALSE}
# detect a resolution column
res_col <- grep("resolution", names(dat_final), ignore.case = TRUE, value = TRUE)
if (length(res_col) > 1) res_col <- res_col[1]

parse_resolution <- function(x) {
  # Accept forms like "1920x1080", "1920 x 1080", "1080 × 1920", etc.
  # Extract the first two integer numbers; if one number, return NA
  nums <- as.integer(stringr::str_extract_all(x, "[0-9]+")[[1]])
  if (length(nums) >= 2) {
    # We don't know orientation; use min as height, max as width
    h <- min(nums[1], nums[2], na.rm = TRUE)
    w <- max(nums[1], nums[2], na.rm = TRUE)
    c(width = w, height = h, area = w*h)
  } else {
    c(width = NA_integer_, height = NA_integer_, area = NA_integer_)
  }
}

if (length(res_col)) {
  res_mat <- t(vapply(dat_final[[res_col]], function(v) {
    if (is.na(v) || !nzchar(v)) return(c(width=NA_integer_, height=NA_integer_, area=NA_integer_))
    parse_resolution(v)
  }, FUN.VALUE = c(width=0L, height=0L, area=0L)))

  dat_final$res_width  <- as.integer(res_mat[, "width"])
  dat_final$res_height <- as.integer(res_mat[, "height"])
  dat_final$res_area   <- as.integer(res_mat[, "area"])
  message("Resolution covariates added: res_height, res_area (and res_width).")
} else {
  message("No resolution column found; skipping resolution covariates.")
}
```

# Quick summaries (optional)
```{r summaries, echo=TRUE, message=TRUE}
# attention status
table(is.na(dat_attn$exclude_reason), useNA = "ifany")

# longstring distribution
summary(dat_attn$longstring_run)
```

# (Placeholder) Modeling examples
Below are example models you can enable later (commented out), once you confirm variable names.

```{r models, eval=FALSE}
# Example: ANOVA without covariates on a score variable 'score'
# dat_model <- dat_final
# dat_model$hierarchy <- factor(dat_model$hierarchy, ordered = FALSE)
# dat_model$chartjunk <- factor(dat_model$chartjunk, ordered = TRUE)
# m1 <- aov(score ~ hierarchy * chartjunk, data = dat_model)
# sjPlot::tab_model(m1)
```


# ---- Appendix: Study 2 Analyses (ANOVA / ANCOVA, sjPlot outputs) ----

```{r setup-appendix, message=FALSE, warning=FALSE}
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
pacman::p_load(dplyr, tidyr, stringr, janitor, ggplot2,
               car, effectsize, sjPlot, emmeans, broom, purrr, tibble)
theme_set(sjPlot::theme_sjplot())
options(contrasts = c("contr.treatment", "contr.poly"))
```

## Data prep (robust to missing flags / columns)

```{r data-prep-appendix}
# prefer attention-filtered data if present
# ---- Resolution covariates (safe; works with separate or combined fields) ----
nm <- names(dat0)

# Try a combined "resolution" field first (e.g., "1920x1080", "1920 x 1080", "1080 × 1920")
res_combined <- nm[stringr::str_detect(nm, regex("^\\s*resolution\\s*$", ignore_case = TRUE))][1] %||% NA_character_

# Try separate width/height fields
cand_h <- nm[stringr::str_detect(nm, regex("(screen|browser|client|res(olution)?)\\s*.*height", ignore_case = TRUE))]
cand_w <- nm[stringr::str_detect(nm, regex("(screen|browser|client|res(olution)?)\\s*.*width",  ignore_case = TRUE))]
res_h_col <- cand_h[1] %||% NA_character_
res_w_col <- cand_w[1] %||% NA_character_

# Initialize result columns if absent
if (!"res_height" %in% names(dat0)) dat0$res_height <- NA_real_
if (!"res_width"  %in% names(dat0)) dat0$res_width  <- NA_real_
if (!"res_area"   %in% names(dat0)) dat0$res_area   <- NA_real_

# 1) Parse combined resolution if available
if (!is.na(res_combined)) {
  nums <- stringr::str_extract_all(dat0[[res_combined]], "[0-9]+")
  got2 <- vapply(nums, function(v) length(v) >= 2, logical(1))
  if (any(got2)) {
    # Use the first two numbers; orientation unknown ⇒ min = height, max = width
    first2 <- lapply(nums, function(v) as.numeric(v[seq_len(min(2, length(v)))]))
    hw <- t(vapply(first2, function(v) {
      if (length(v) < 2 || anyNA(v)) return(c(h=NA_real_, w=NA_real_))
      c(h = min(v[1], v[2], na.rm = TRUE), w = max(v[1], v[2], na.rm = TRUE))
    }, FUN.VALUE = c(h=NA_real_, w=NA_real_)))
    # Only fill missing values; don't overwrite existing non-NA
    dat0$res_height <- ifelse(is.na(dat0$res_height), hw[, "h"], dat0$res_height)
    dat0$res_width  <- ifelse(is.na(dat0$res_width),  hw[, "w"], dat0$res_width)
  }
}

# 2) Use separate height/width columns if present
if (!is.na(res_h_col)) {
  dat0$res_height <- ifelse(
    is.na(dat0$res_height),
    suppressWarnings(as.numeric(dat0[[res_h_col]])),
    dat0$res_height
  )
}
if (!is.na(res_w_col)) {
  dat0$res_width <- ifelse(
    is.na(dat0$res_width),
    suppressWarnings(as.numeric(dat0[[res_w_col]])),
    dat0$res_width
  )
}

# 3) Compute area if either pair is available
dat0$res_area <- ifelse(
  is.na(dat0$res_area) & is.finite(dat0$res_height) & is.finite(dat0$res_width),
  dat0$res_height * dat0$res_width,
  dat0$res_area
)

# Quick peek (optional)
# summary(dat0$res_height); summary(dat0$res_width); summary(dat0$res_area)


# Coerce score to numeric if needed (Qualtrics scores often come in as character)
score_col <- c("Score","score","SC0","Q_Score","TotalScore")
score_col <- score_col[score_col %in% names(dat0)][1] %||% "Score"
if (!score_col %in% names(dat0)) stop("Could not find a score column. Expected one of: ", paste(score_col, collapse=", "))

dat0 <- dat0 %>% mutate(score = suppressWarnings(as.numeric(.data[[score_col]])))

# If hierarchy / chartjunk not present, derive from version
if (!all(c("hierarchy","chartjunk") %in% names(dat0))) {
  version_col <- c("version","Version","condition","Condition")
  version_col <- version_col[version_col %in% names(dat0)][1] %||% "version"
  if (!version_col %in% names(dat0)) stop("Need either (hierarchy, chartjunk) or a version column to derive from.")

  dat0 <- dat0 %>% mutate(
    hierarchy = case_when(
      str_detect(.data[[version_col]], regex("NoHierarchy", ignore_case = TRUE)) ~ "NoHierarchy",
      str_detect(.data[[version_col]], regex("RoomHierarchy", ignore_case = TRUE)) ~ "RoomHierarchy",
      str_detect(.data[[version_col]], regex("ImportanceHierarchy", ignore_case = TRUE)) ~ "ImportanceHierarchy",
      str_detect(.data[[version_col]], regex("FloorplanHierarchy", ignore_case = TRUE)) ~ "FloorplanHierarchy",
      TRUE ~ NA_character_
    ),
    chartjunk = case_when(
      str_detect(.data[[version_col]], regex("HighChartjunk|ChartjunkHigh|High", ignore_case = TRUE)) ~ "High",
      str_detect(.data[[version_col]], regex("MediumChartjunk|ChartjunkMedium|Medium", ignore_case = TRUE)) ~ "Medium",
      str_detect(.data[[version_col]], regex("LowChartjunk|ChartjunkLow|Low", ignore_case = TRUE)) ~ "Low",
      TRUE ~ NA_character_
    )
  )
}

# Order/format factors
# hierarchy is nominal (not ordinal)
dat0 <- dat0 %>% mutate(
  hierarchy = factor(hierarchy),
  # chartjunk is ordinal
  chartjunk = factor(chartjunk,
                     levels = c("Low","Medium","High"),
                     ordered = TRUE)
)

# Longstring exclusion if available
if ("longstring_flag" %in% names(dat0)) {
  dat0 <- dat0 %>% filter(!coalesce(longstring_flag, FALSE))
}

# SurveyWave optional (used only for covariate/control if present)
if ("surveywave" %in% names(dat0)) {
  dat0 <- dat0 %>% mutate(surveywave = factor(surveywave))
}

# Resolution covariates (robust detection)
nm <- names(dat0)

cand_h <- nm[str_detect(nm, regex("(screen|browser|res(olution)?|client).*height", ignore_case = TRUE))]
cand_w <- nm[str_detect(nm, regex("(screen|browser|res(olution)?|client).*width",  ignore_case = TRUE))]

res_h_col <- cand_h[1] %||% NA_character_
res_w_col <- cand_w[1] %||% NA_character_

# Explicit area column, if any
cand_area <- nm[str_detect(nm, regex("res(olution)?_?area|screen_?area|area_px", ignore_case = TRUE))]
res_area_col <- cand_area[1] %||% NA_character_

dat0 <- dat0 %>% mutate(
  res_height = suppressWarnings(as.numeric(.data[[res_h_col]])),
  res_width  = suppressWarnings(as.numeric(.data[[res_w_col]]))
)

# Compute area if we can, else use provided
if (!is.na(res_area_col)) {
  dat0 <- dat0 %>% mutate(res_area = suppressWarnings(as.numeric(.data[[res_area_col]])))
} else if (all(c("res_width","res_height") %in% names(dat0))) {
  dat0 <- dat0 %>% mutate(res_area = res_width * res_height)
}

# Keep analysis-ready subset
datA <- dat0 %>%
  filter(!is.na(score), !is.na(hierarchy), !is.na(chartjunk)) %>%
  rename(chartjunk_ord = chartjunk) %>%
  droplevels()

summary(datA$hierarchy); summary(datA$chartjunk_ord)
```

## ANOVA: hierarchy × chartjunk (no covariates)

```{r anova-main, fig.width=7, fig.height=4}
m_anova <- lm(score ~ hierarchy * chartjunk_ord, data = datA)
anova(m_anova)

# sjPlot table of coefficients
sjPlot::tab_model(m_anova, show.std = TRUE, show.aic = TRUE, dv.labels = "Score")

# Interaction plot
sjPlot::plot_model(m_anova, type = "int", terms = c("hierarchy", "chartjunk_ord"))
```

### Main-effect plots (predicted means)

```{r main-effects, fig.width=6.5, fig.height=4}
# hierarchy main effect (marginalized over chartjunk)
sjPlot::plot_model(m_anova, type = "pred", terms = "hierarchy")

# chartjunk (ordinal) main effect
sjPlot::plot_model(m_anova, type = "pred", terms = "chartjunk_ord")
```

### Effect sizes (η² / partial η²)

```{r anova-effectsizes}
anova_tbl <- anova(m_anova)
effectsize::eta_squared(anova_tbl, partial = FALSE)
effectsize::eta_squared(m_anova, partial = TRUE)  # partial η²
```

## ANCOVA: adjusting for resolution (height vs. area)

```{r ancova-resolution, fig.width=7, fig.height=4}
# Choose available resolution covariates
has_h <- "res_height" %in% names(datA) && any(is.finite(datA$res_height))
has_a <- "res_area"   %in% names(datA) && any(is.finite(datA$res_area))

models <- list()

if (has_h) {
  m_ancova_h <- lm(score ~ hierarchy * chartjunk_ord + res_height + surveywave, data = datA)
  models$Height <- m_ancova_h
}
if (has_a) {
  m_ancova_a <- lm(score ~ hierarchy * chartjunk_ord + res_area + surveywave, data = datA)
  models$Area <- m_ancova_a
}

if (length(models) == 0) stop("No usable resolution covariate found (height or area).")

# sjPlot tables
do.call(sjPlot::tab_model, c(models, list(show.std = TRUE, show.aic = TRUE, dv.labels = "Score")))

# Compare which resolution covariate is more informative
cmp <- tibble::tibble(
  Model = names(models),
  AIC   = purrr::map_dbl(models, AIC),
  BIC   = purrr::map_dbl(models, BIC),
  R2    = purrr::map_dbl(models, ~ broom::glance(.x)$r.squared %||% NA_real_),
  AdjR2 = purrr::map_dbl(models, ~ broom::glance(.x)$adj.r.squared %||% NA_real_)
) %>% arrange(AIC)
cmp
```

### ANCOVA interaction & main-effect plots

```{r ancova-plots, fig.width=7, fig.height=4, results='hide'}
m_ref <- models[[1]]

# Interaction (hierarchy × chartjunk)
sjPlot::plot_model(m_ref, type = "int", terms = c("hierarchy", "chartjunk_ord"))

# Main effects
sjPlot::plot_model(m_ref, type = "pred", terms = "hierarchy")
sjPlot::plot_model(m_ref, type = "pred", terms = "chartjunk_ord")
```

### Trend test for chartjunk (ordinal)

```{r chartjunk-trend}
# Polynomial (linear, quadratic) trends via emmeans
emms <- emmeans::emmeans(m_anova, ~ chartjunk_ord)
emmeans::contrast(emms, method = "poly")
```

### Optional: report-friendly tables

```{r report-tables, results='asis'}
# Type-III ANOVA (if you prefer)
car::Anova(m_anova, type = 3)

# Partial η² for ANCOVA (first model shown)
m_ref <- models[[1]]
effectsize::eta_squared(car::Anova(m_ref, type = 3), partial = TRUE)
```
