---
title: "Study 2 — Scoring, Attention Filter, Effects (future‑ready)"
output:
  html_document:
    df_print: paged
---


```{r setup, message=FALSE, warning=FALSE}
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
pacman::p_load(readr, dplyr, stringr, janitor, ggplot2, sjPlot, car, effectsize, emmeans, broom)

# ---------- Set your data path here ----------
# Use one of: forward slashes / double backslashes / raw string
# path <- "D:/Dropbox/USD/Dissertation/Collab folder/Data/Dissertation+-+Study+2_September+29,+2025_21.58.csv"

# 3) R 4.0+ raw string
path <- r"(D:\Dropbox\USD\Dissertation\Collab folder\Data\Dissertation+-+Study+2_September+30,+2025_10.01.csv)"

```

## Load + clean Qualtrics export (keep header row 2; drop ImportId row)

```{r load_clean}
# 1) Row 1 are labels; make Row 2 the header
raw <- readr::read_csv(
  path,
  skip = 1,
  col_types = readr::cols(.default = readr::col_character()),
  show_col_types = FALSE
)

# 2) Remove the label/metadata row (the original row 3) if present
dat <- raw %>%
  dplyr::filter(!dplyr::if_any(dplyr::everything(), ~ grepl("\\bImportId\\b", .x, ignore.case = TRUE)))

# Optional: clean names for convenience in parallel object
dat_clean <- dat %>% janitor::clean_names()

dat = dat %>%
  filter(dat$`Response Type` != "Survey Preview")

# Peek
dplyr::glimpse(dat)
```

## Attention check → remove careless responders

We drop anyone who **did not** answer “Smart Home” to **“What type of dashboard did you see?”**

```{r attention_filter}
# Find the attention-check column by name (case-insensitive; optional trailing '?')
attn_col <- grep("^What type of dashboard did you see\\??$", names(dat), ignore.case = TRUE, value = TRUE)

if (length(attn_col) == 0) {
  message("Attention check column not found — no rows removed at this step.")
  dat_attn <- dat
} else {
  # Keep rows that say 'smart home' (case-insensitive, allow whitespace variance)
  keep_idx <- grepl("smart\\s*home", dat[[attn_col]], ignore.case = TRUE)
  removed_n <- sum(!keep_idx, na.rm = TRUE)
  message("Attention filter removed: ", removed_n, " row(s).")
  dat_attn <- dat[keep_idx %in% TRUE, , drop = FALSE]
}

# If you want to also filter speeders (optional; uncomment and set a threshold)
# dur_col <- grep("^Duration.*seconds\\)?$", names(dat_attn), ignore.case = TRUE, value = TRUE)
# if (length(dur_col) == 1) {
#   thr <- quantile(suppressWarnings(as.numeric(dat_attn[[dur_col]])), probs = 0.05, na.rm = TRUE)
#   dat_attn <- dat_attn %>% dplyr::filter(suppressWarnings(as.numeric(.data[[dur_col]])) >= thr)
# }
```

## Factors: chartjunk (ordinal) and hierarchy (nominal)

```{r factors}
dat_attn <- dat_attn %>% dplyr::mutate(version = as.character(.data[["Version"]] %||% .data[["version"]]))

# Extract chartjunk (High/Medium/Low) and set ordered
cj <- stringr::str_match(dat_attn$version, "(High|Medium|Low)Chartjunk$")[,2]
dat_attn$chartjunk <- factor(cj, levels = c("Low","Medium","High"), ordered = TRUE)

# Extract hierarchy by removing the tail
hier_raw <- stringr::str_remove(dat_attn$version, "(High|Medium|Low)Chartjunk$")
dat_attn$hierarchy <- dplyr::recode(hier_raw,
  "NoHierarchy" = "none",
  "FloorplanHierarchy" = "floorplan",
  "RoomHierarchy" = "room",
  "ImportanceHierarchy" = "importance",
  .default = NA_character_
) |> factor()

# Sanity
table(dat_attn$hierarchy, dat_attn$chartjunk, useNA = "ifany")
```

## Outcome: Score (use existing if present; otherwise from SC* columns)

```{r outcome}
# Prefer an existing 'Score' column if present
score_col <- grep("^Score$", names(dat_attn), value = TRUE)
if (length(score_col) == 1) {
  dat_attn$Total_Score <- suppressWarnings(as.numeric(dat_attn[[score_col]]))
  source_of_score <- "existing Score column"
} else {
  # Fall back to Qualtrics scoring categories SC0, SC1, ...
  sc_cols <- grep("^SC\\d+$", names(dat_attn), value = TRUE)
  if (length(sc_cols) == 0) {
    warning("No Score column and no SC* columns found; Total_Score set to NA.")
    dat_attn$Total_Score <- NA_real_
    source_of_score <- "none"
  } else {
    dat_attn[sc_cols] <- lapply(dat_attn[sc_cols], function(x) suppressWarnings(as.numeric(x)))
    dat_attn$Total_Score <- rowSums(dat_attn[sc_cols], na.rm = TRUE)
    source_of_score <- "SC* sum"
  }
}
message("Total_Score source: ", source_of_score)
dat_attn$Percent_Score <- dat_attn$Total_Score / max(dat_attn$Total_Score, na.rm = TRUE) * 100
summary(dat_attn[c("Total_Score","Percent_Score")])
```

```{r}
# Packages
if (!requireNamespace("sjPlot", quietly = TRUE)) install.packages("sjPlot")
if (!requireNamespace("car", quietly = TRUE)) install.packages("car")
library(sjPlot); library(car)

# --- Data choice ---
if (!exists("dat_attn")) stop("dat_attn not found. Build it first (post-attention-filter data).")

# --- Detect columns ---
score_var <- if ("Score" %in% names(dat_attn)) "Score" else
             if ("Total_Score" %in% names(dat_attn)) "Total_Score" else
             stop("No Score or Total_Score column found in dat_attn.")

cj_var <- if ("chartjunk" %in% names(dat_attn)) "chartjunk" else
          if ("charjunk" %in% names(dat_attn)) "charjunk" else
          stop("No chartjunk/charjunk column found in dat_attn.")

h_var <- if ("hierarchy" %in% names(dat_attn)) "hierarchy" else
         stop("No hierarchy column found in dat_attn.")

# --- Ensure factor types ---
dat_attn[[h_var]]  <- as.factor(dat_attn[[h_var]])

# Enforce Low < Medium < High if present; otherwise make ordered as-is
if (all(c("Low","Medium","High") %in% unique(na.omit(dat_attn[[cj_var]])))) {
  dat_attn[[cj_var]] <- factor(dat_attn[[cj_var]], levels = c("Low","Medium","High"), ordered = TRUE)
} else {
  dat_attn[[cj_var]] <- factor(dat_attn[[cj_var]], ordered = TRUE)
}

# Polynomial contrasts for ordered factor
if (is.ordered(dat_attn[[cj_var]])) {
  contrasts(dat_attn[[cj_var]]) <- contr.poly(length(levels(dat_attn[[cj_var]])))
}

# --- Fit model (no covariates) ---
form <- as.formula(paste0(score_var, " ~ ", h_var, " * ", cj_var))
cat("Model formula:", deparse(form), "\n")
lm_model <- lm(form, data = dat_attn, na.action = na.exclude)

# --- ANOVA (Type II) ---
anova2 <- car::Anova(lm_model, type = 2)
print(anova2)

# --- sjPlot outputs ---
# Coefficient table
tab_model(lm_model,
          title = "Score ~ Hierarchy * Chartjunk (no covariates)",
          show.ci = FALSE, show.se = TRUE, show.stat = TRUE, show.p = TRUE,
          dv.labels = score_var)

# Interaction plot
plot_model(lm_model, type = "int", terms = c(h_var, cj_var))
# Optional marginal effects:
# plot_model(lm_model, type = "eff", terms = h_var)
# plot_model(lm_model, type = "eff", terms = cj_var)


```


## Covariates: resolution + the 4 post‑Q20 items (use if present)

```{r covariates}
# Explicit covariates by their survey question text
covariates <- c(
  "How often do you use data dashboards in your life?",
  "How often do you use a smart home dashboard in your life?",
  "How clear was the dashboard’s visual design (e.g., text readability, contrast, spacing)?",
  "How clear was the dashboard's informational organization (grouping, labels, layout)?"
)

# Present-in-data detection supports raw & cleaned names
present <- intersect(c(covariates, janitor::make_clean_names(covariates)), names(dat_attn))

# Optional: coerce some types if found
for (nm in present) {
  # If it's a Likert from 1-5 etc., coerce to ordered numeric carefully
  # Here we try numeric-first, otherwise leave as factor
  suppressWarnings({
    asnum <- as.numeric(dat_attn[[nm]])
    if (!all(is.na(asnum))) dat_attn[[nm]] <- asnum
  })
}

# resolution covariate
res_col <- grep("^resolution$", names(dat_attn), ignore.case = TRUE, value = TRUE)
covar_vec <- c(present, res_col)
covar_vec <- unique(covar_vec[covar_vec %in% names(dat_attn)])

cat("Covariates included:", paste(covar_vec, collapse = ", "), "\n")
```

## Model (future‑ready): `Total_Score ~ hierarchy * chartjunk + covariates`

```{r model, fig.height=4.5, fig.width=7, message=FALSE, warning=FALSE}
# Helper to backtick names safely
bt <- function(x) paste0("`", gsub("`","``", x), "`")

# Build formula with interaction + covariates
rhs <- c("hierarchy * chartjunk", if (length(covar_vec)) bt(covar_vec) else NULL)
form <- as.formula(paste("Total_Score ~", paste(rhs, collapse = " + ")))
cat("Model formula:", deparse(form), "\n")

# Fit
lm_model <- lm(form, data = dat_attn)

# Type II ANOVA (common with covariates)
anova2 <- car::Anova(lm_model, type = 2)
anova2
```

## Effect sizes (η², partial η², ω²)

```{r effects}
# From the fitted model
eta <- effectsize::eta_squared(lm_model, partial = FALSE, ci = 0.90)
eta_p <- effectsize::eta_squared(lm_model, partial = TRUE, ci = 0.90)
omega <- effectsize::omega_squared(lm_model, partial = FALSE, ci = 0.90)

list(
  eta_squared = eta,
  partial_eta_squared = eta_p,
  omega_squared = omega
)
```

## Estimated marginal means (optional) and interaction plot

```{r emmeans_plot, fig.height=4.5, fig.width=7, message=FALSE, warning=FALSE}
# Only if factors exist
if (all(c("hierarchy","chartjunk") %in% names(dat_attn))) {
  sjPlot::plot_model(lm_model, type = "int", terms = c("hierarchy","chartjunk"))
}
```

## Cell means

```{r cell_means}
cell_means <- dat_attn %>%
  dplyr::group_by(hierarchy, chartjunk) %>%
  dplyr::summarise(
    n = dplyr::n(),
    mean_total = mean(Total_Score, na.rm = TRUE),
    sd_total = stats::sd(Total_Score, na.rm = TRUE),
    .groups = "drop"
  )
cell_means
```

## QA: exclusions + missingness report

```{r qa}
# Count exclusions (attention)
total_n <- nrow(dat)
kept_n  <- nrow(dat_attn)
cat("Total rows:", total_n, " | Kept after attention filter:", kept_n, " | Removed:", total_n - kept_n, "\n")

# Missingness in covariates used
if (length(covar_vec)) {
  miss <- dat_attn %>%
    dplyr::summarise(across(all_of(covar_vec), ~ mean(is.na(.x), na.rm = TRUE))) %>%
    tidyr::pivot_longer(dplyr::everything(), names_to = "var", values_to = "pct_missing")
  miss
}
```
