---
title: "Study 2 — Positions FIXED (QSF keys, robust DO parsing)"
output:
  html_document:
    df_print: paged
---

```{r setup, message=FALSE, warning=FALSE}
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
pacman::p_load(readr, dplyr, stringr, janitor, ggplot2, jsonlite, car, effectsize, sjPlot, vcd, rstatix, tidyr, purrr, tibble)
`%||%` <- function(a,b) if (is.null(a) || length(a)==0) b else a
```

## Paths

```{r paths}
path <- r"(D:\Dropbox\USD\Dissertation\Collab folder\Data\Dissertation+-+Study+2_September+30,+2025_10.01.csv)"
csv_path <- r"(D:/Dropbox/USD/Dissertation/Collab folder/Data/Dissertation+-+Study+2_September+30,+2025_10.01.csv)"
qsf_path <- r"(D:/Dropbox/USD/Dissertation/Collab folder/Data/Dissertation_-_Study_2.qsf)"
```

## Load & clean CSV (row 2 headers; drop ImportId row)

```{r load_csv}
raw <- readr::read_csv(csv_path, skip = 1, col_types = readr::cols(.default = readr::col_character()), show_col_types = FALSE)
dat <- raw %>%
  dplyr::filter(!dplyr::if_any(dplyr::everything(),
                               ~ grepl("ImportId", .x, fixed = TRUE)))

dat_orig <- dat
```

## Parse QSF → Q1–Q20 (tags, QIDs, choices, default order, randomization)

```{r parse_qsf}
stopifnot(file.exists(qsf_path))
qsf <- jsonlite::fromJSON(qsf_path, simplifyVector = FALSE)
sq  <- Filter(function(el) identical(el$Element, "SQ"), qsf$SurveyElements)

get_tag   <- function(p) p$DataExportTag %||% NA_character_
get_qid   <- function(p) p$QuestionID    %||% NA_character_
get_text  <- function(p) p$QuestionText  %||% NA_character_
get_rand  <- function(p) { ra <- p$Randomization %||% list(); if (isTRUE(ra$Advanced) || isTRUE(ra$RandomizeAnswers) || !is.null(ra$Type)) TRUE else FALSE }
get_choice_order <- function(p) { ord <- p$ChoiceOrder %||% names(p$Choices); unname(as.character(ord)) }
get_choices_tbl <- function(p) {
  ch <- p$Choices %||% list()
  keys <- names(ch) %||% character(0)
  if (!length(keys)) return(tibble::tibble(key=character(0), text=character(0), recode=character(0)))
  rv <- p$RecodeValues %||% list()
  recode_vec <- vapply(keys, function(k) as.character(rv[[k]] %||% NA_character_), character(1))
  tibble::tibble(key = as.character(keys), text = vapply(ch, function(x) x$Display %||% "", character(1)), recode = recode_vec)
}
q_meta <- lapply(sq, function(el) {
  p <- el$Payload
  tibble::tibble(tag=get_tag(p), qid=get_qid(p), text=get_text(p), randomize=get_rand(p),
                 choice_order_default=list(get_choice_order(p)), choices=list(get_choices_tbl(p)))
}) %>% dplyr::bind_rows() %>% dplyr::filter(!is.na(tag), grepl("^Q([1-9]|1[0-9]|20)$", tag)) %>%
     dplyr::arrange(factor(tag, levels = paste0("Q", 1:20)))
q_meta
```

## Map Q1–Q20 to CSV columns and build per‑QID **position** columns

```{r positions, message=FALSE, warning=FALSE}
find_csv_col <- function(q_text, cols){
  m <- which(cols == q_text)[1]; if (!is.na(m)) return(cols[m])
  m <- which(startsWith(cols, q_text))[1]; if (!is.na(m)) return(cols[m])
  NA_character_
}
cols <- names(dat)

q_meta <- q_meta %>% dplyr::mutate(
  csv_col       = vapply(text, find_csv_col, character(1), cols = cols),
  csv_col_order = ifelse(is.na(csv_col), NA_character_, paste0(csv_col, " - Display Order"))
)

# Safe regex helpers (POSIX classes)
answer_to_key <- function(answer_val, choices){
  if (is.na(answer_val) || !length(choices$key)) return(NA_character_)
  ans <- as.character(answer_val)

  # 1) exact text match
  idx <- match(ans, choices$text)
  if (!is.na(idx)) return(choices$key[idx])

  # 2) recode match (when export stores numeric recode)
  idx <- match(ans, choices$recode)
  if (!is.na(idx)) return(choices$key[idx])

  # 3) "1 - Choice text" style: extract leading digits (POSIX)
  lead <- regmatches(ans, regexpr("^[[:space:]]*[0-9]+", ans))
  if (length(lead) && nchar(lead)) {
    lead <- sub("^[[:space:]]*([0-9]+).*", "\\1", lead)
    idx <- match(lead, choices$key)
    if (!is.na(idx)) return(choices$key[idx])
  }

  # 4) already a key?
  if (ans %in% choices$key) return(ans)

  NA_character_
}

order_to_keys <- function(order_str, choices, default_keys){
  if (is.na(order_str) || !nzchar(order_str)) return(default_keys)

  # Detect delimiter
  delim <- if (grepl("\\|", order_str)) "\\|" else if (grepl(";", order_str)) ";" else ","
  toks <- trimws(unlist(strsplit(order_str, delim)))
  if (!length(toks)) return(default_keys)

  # A) tokens are keys already
  if (all(toks %in% choices$key)) return(toks)

  # B) tokens are full texts
  if (all(toks %in% choices$text)) {
    return(choices$key[match(toks, choices$text)])
  }

  # C) tokens like "1 - Choice text": extract leading digits with POSIX
  lead <- sub("^[[:space:]]*([0-9]+).*", "\\1", toks)
  if (all(lead %in% choices$key)) return(lead)

  # fallback
  default_keys
}


pos_map <- list()
for (i in seq_len(nrow(q_meta))) {
  meta <- q_meta[i, ]; qcol <- meta$csv_col; if (is.na(qcol)) next
  odcol <- meta$csv_col_order; choices <- meta$choices[[1]]; def_keys <- meta$choice_order_default[[1]]
  qid <- meta$qid; tag <- meta$tag; pos_name <- paste0(tag, "_", qid, "_position")
  ans_keys <- vapply(dat[[qcol]], answer_to_key, character(1), choices = choices)
ord_keys <- if (!is.na(odcol) && odcol %in% names(dat)) {
  lapply(dat[[odcol]], order_to_keys, choices = choices, default_keys = def_keys)
} else {
  rep(list(def_keys), nrow(dat))
}

  pos <- mapply(function(k, ord) { if (is.na(k) || !length(ord)) return(NA_integer_); p <- match(k, ord); if (is.na(p)) NA_integer_ else as.integer(p) }, ans_keys, ord_keys)
  dat[[pos_name]] <- as.integer(pos)
  pos_map[[pos_name]] <- pos_name
}

tags <- unique(q_meta$tag)
for (tg in tags) {
  cols_tag <- grep(paste0("^", tg, "_.*_position$"), names(dat), value = TRUE)
  if (length(cols_tag)) dat[[paste0(tg, "_position")]] <- do.call(dplyr::coalesce, dat[cols_tag])
}

item_cols_Q1_Q20 <- q_meta$csv_col
pos_cols_Q1_Q20  <- paste0(q_meta$tag, "_position")
list(item_cols = item_cols_Q1_Q20, pos_cols  = pos_cols_Q1_Q20) %>% print()
```

## Attention filter (Smart Home)

```{r attention_filter, message=FALSE, warning=FALSE}
attn_col <- grep("^What type of dashboard did you see\\??$", names(dat),
                 ignore.case = TRUE, value = TRUE)

if (length(attn_col) == 0) { dat_attn <- dat; dat_attn$exclude_reason <- NA_character_
} else { keep_idx <- grepl("smart\\s*home", dat[[attn_col]], ignore.case = TRUE)
         dat$exclude_reason <- ifelse(keep_idx, NA_character_, "attention_fail")
         dat_attn <- dat[keep_idx %in% TRUE, , drop = FALSE] }
```

## Longstring (position‑aware; Q1–Q20)

```{r longstring_positions}
present_pos <- pos_cols_Q1_Q20[pos_cols_Q1_Q20 %in% names(dat_attn)]
if (!length(present_pos)) { dat_final <- dat_attn
} else {
  vals <- dat_attn[, present_pos, drop = FALSE]
  ls_run <- apply(vals, 1, function(row) { v <- as.integer(row); r <- rle(ifelse(is.na(v), paste0("NA_", seq_along(v)), paste0("V", v))); max(r$lengths, na.rm = TRUE) })
  q3  <- stats::quantile(ls_run, 0.75, na.rm = TRUE); iqr <- stats::IQR(ls_run, na.rm = TRUE); thr <- q3 + 1.5 * iqr
  flag_ls <- ls_run > thr
  dat_attn$longstring_run <- ls_run
  dat_attn$exclude_reason <- ifelse(is.na(dat_attn$exclude_reason) & flag_ls, "longstring_outlier",
                             ifelse(!is.na(dat_attn$exclude_reason) & flag_ls, paste(dat_attn$exclude_reason, "+ longstring_outlier"), dat_attn$exclude_reason))
  dat_final <- dat_attn[!flag_ls, , drop = FALSE]
  cat("Position-based longstring: threshold >", round(thr,2), " | removed", sum(flag_ls, na.rm = TRUE), "rows.
")
}
```

## Resolution → res_height & res_area

```{r resolution_parse, message=FALSE, warning=FALSE}
res_col <- grep("^resolution$", names(dat_final), ignore.case = TRUE, value = TRUE)
if (length(res_col) == 1) {
  res_str <- dat_final[[res_col]]; wh <- do.call(rbind, strsplit(as.character(res_str), "[^0-9]+"))
  suppressWarnings({ res_width  <- as.numeric(wh[,1]); res_height <- as.numeric(wh[,2]) })
  dat_final$res_width  <- res_width; dat_final$res_height <- res_height; dat_final$res_area <- res_width * res_height
} else { message("No 'resolution' field detected; res_width/height/area not added.") }
```

## Factors & outcome

```{r factors_outcome}
dat_final <- dat_final %>% dplyr::mutate(version = as.character(.data[["Version"]] %||% .data[["version"]]))
cj <- stringr::str_match(dat_final$version, "(High|Medium|Low)Chartjunk$")[,2]
dat_final$chartjunk <- factor(cj, levels = c("Low","Medium","High"), ordered = TRUE)
hier_raw <- stringr::str_remove(dat_final$version, "(High|Medium|Low)Chartjunk$")
dat_final$hierarchy <- dplyr::recode(hier_raw, "NoHierarchy"="none","FloorplanHierarchy"="floorplan","RoomHierarchy"="room","ImportanceHierarchy"="importance", .default = NA_character_) |> factor()
score_col <- grep("^Score$", names(dat_final), value = TRUE)
if (length(score_col) == 1) { dat_final$Total_Score <- suppressWarnings(as.numeric(dat_final[[score_col]]))
} else { sc_cols <- grep("^SC\\d+$", names(dat_final), value = TRUE)
         if (length(sc_cols)) { dat_final[sc_cols] <- lapply(dat_final[sc_cols], function(x) suppressWarnings(as.numeric(x))); dat_final$Total_Score <- rowSums(dat_final[sc_cols], na.rm = TRUE) }
         else dat_final$Total_Score <- NA_real_ }
```

## Model (hierarchy × chartjunk + resolution covariates)

```{r model, message=FALSE, warning=FALSE}
covar_vec <- intersect(c("res_height","res_area"), names(dat_final))
if (is.ordered(dat_final$chartjunk)) contrasts(dat_final$chartjunk) <- contr.poly(length(levels(dat_final$chartjunk)))
rhs <- c("hierarchy * chartjunk", if (length(covar_vec)) covar_vec else NULL)
form <- as.formula(paste("Total_Score ~", paste(rhs, collapse = " + "))); cat("Model formula:", deparse(form), "
")
lm_model <- lm(form, data = dat_final); car::Anova(lm_model, type = 2)
```

## Effect sizes

```{r effects}
list(eta_sq=effectsize::eta_squared(lm_model, partial = FALSE, ci = 0.90),
     partial_eta_sq=effectsize::eta_squared(lm_model, partial = TRUE, ci = 0.90),
     omega_sq=effectsize::omega_squared(lm_model, partial = FALSE, ci = 0.90))
```

## Item ↔ hierarchy diagnostics (response & position)

```{r item_diag, message=FALSE, warning=FALSE, fig.height=5, fig.width=8}
item_cols <- item_cols_Q1_Q20[!is.na(item_cols_Q1_Q20)]
resp_results <- list()
for (nm in item_cols) {
  x <- dat_final[[nm]]; g <- dat_final$hierarchy; x_num <- suppressWarnings(as.numeric(x)); uniq <- length(unique(na.omit(x)))
  if (!all(is.na(x_num)) && uniq <= 7) { kt <- kruskal.test(x_num ~ g); H <- unname(kt$statistic); k <- nlevels(g); n <- sum(!is.na(x_num) & !is.na(g)); eps2 <- (H - k + 1) / (n - k); resp_results[[nm]] <- data.frame(item = nm, method = "Kruskal-epsilon^2", effect = eps2, p = kt$p.value) }
  else { tbl <- table(as.factor(x), g, useNA = "no"); chi <- suppressWarnings(chisq.test(tbl)); n <- sum(tbl); r <- nrow(tbl); c <- ncol(tbl); V <- sqrt(unname(chi$statistic) / (n * (min(r - 1, c - 1)))); resp_results[[nm]] <- data.frame(item = nm, method = "CramersV", effect = V, p = chi$p.value) }
}
resp_fx <- dplyr::bind_rows(resp_results)

pos_cols <- pos_cols_Q1_Q20[pos_cols_Q1_Q20 %in% names(dat_final)]
pos_results <- list()
for (nm in pos_cols) {
  pos <- dat_final[[nm]]; g <- dat_final$hierarchy; if (is.null(pos) || all(is.na(pos))) next
  kt <- kruskal.test(pos ~ g); H <- unname(kt$statistic); k <- nlevels(g); n <- sum(!is.na(pos) & !is.na(g)); eps2 <- (H - k + 1) / (n - k)
  pos_results[[nm]] <- data.frame(item = nm, method = "Position (Kruskal ε²)", effect = eps2, p = kt$p.value)
}
pos_fx <- dplyr::bind_rows(pos_results)

all_fx <- dplyr::bind_rows(resp_fx, pos_fx) %>% dplyr::arrange(dplyr::desc(effect)); head(all_fx, 10)

topN <- head(all_fx$item, 15); plot_df <- all_fx %>% dplyr::filter(item %in% topN)
ggplot(plot_df, aes(x = reorder(item, effect), y = effect, fill = method)) + geom_col() + coord_flip() +
  labs(title = "Item–Hierarchy effects (response & position)", x = "Item / Position metric", y = "Effect size") + theme_minimal()
```

## QA — exclusions & balance

```{r qa}
total_n <- nrow(dat); kept_n  <- nrow(dat_final); cat("Total rows:", total_n, " | Kept after filters:", kept_n, " | Removed:", total_n - kept_n, "
")
if ("exclude_reason" %in% names(dat_attn)) { table(dat_attn$exclude_reason, useNA = "ifany") }
if (all(c("hierarchy","chartjunk") %in% names(dat_final))) { with(dat_final, table(hierarchy, chartjunk)) }
if (length(pos_cols_Q1_Q20[pos_cols_Q1_Q20 %in% names(dat_final)]) > 0) {
  pos_long <- tidyr::pivot_longer(dat_final, all_of(pos_cols_Q1_Q20[pos_cols_Q1_Q20 %in% names(dat_final)]), names_to = "item", values_to = "position")
  ggplot(pos_long, aes(x = factor(position))) + geom_bar() + facet_wrap(~ item, scales = "free_y") + theme_minimal() + labs(title = "Chosen positions per item", x = "Position index", y = "Count")
}
```

